### 2024-09-04 Corrections_SAM_SS3_Krill_481.Rmd
adding layaout in correction file

---
title: "Integrated approach to modeling krill population dynamics in Western Antarctic Peninsula. Spatial and ecosystem considerations"
subtitle: "WG-FSA 2024. Working Paper"
author:
  - Mauricio Mardones^[Universidad de Magallanes, mamardon@umag.cl, Milenio Base]
  - Lucas Krüger ^[Instituto Antártico Chileno, Milenio Base]
  - Francisco Santa Cruz ^[Instituto Antártico Chileno, Universidad de Magallanes, Milenio Base]
  - César Cárdenas^[Instituto Antártico Chileno, Milenio Base]
  - George Watters^[Ecosystem Science Division, Southwest Fisheries Science Center, NOAA, USA]
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output:
  bookdown::pdf_document2:
    number_sections: false
fig_caption: yes
bibliography: SA_krill.bib
csl: apa.csl
link-citations: yes
toc: false
linkcolor: blue
linestretch: 1.3
header-includes:
- \fontsize{12}{16}
- \selectfont
- \usepackage{lscape}
---
\newpage
\tableofcontents
\newpage

```{r setup1, echo=FALSE}
rm(list = ls())
knitr::opts_chunk$set(eval=TRUE,
                      echo = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      fig.align = 'center',
                      fig.pos = "H",
                      dev = 'jpeg',
                      dpi = 300,
                      tidy.opts=list(width.cutoff=50), 
                      tidy=TRUE)
```

```{r message=FALSE, eval=TRUE}
library(here)
library(kableExtra)
library(ggpubr)
library(tibble)
library(readxl)
library(openxlsx)
library(r4ss)
dir1<-here("s1")
```

# BACKGROUND

This study addresses key recommendations highlighted in the preliminary report of the Working Group on Statistics, Assessment, and Modelling (WG-SAM-2024/26) [@Mardones2023] regarding the integration of diverse information sources for understanding the population dynamics of Antarctic krill (*Euphausia superba*) within an integrated stock assessment model, particularly in the context of environmental change and fisheries management challenges. The main observations were: 1) Concerns about using a length-based model due to the scarcity of direct age data, as this approach avoids the approximations required when converting length data to age data; 2) The need to consider the impact of changes in trawl designs over time and to reassess the assumption that the Antarctic Peninsula is a closed system; 3) The limitations of using krill length and biological composition data from fishing vessels, given the significant variability and potential biases, and the importance of understanding the interaction between fisheries and krill-dependent predators; 4) A recommendation to present a progression of model implementations from simple to complex to better evaluate model assumptions, investigate patterns in residuals, and assess the realism of parameter values used, while also integrating recent findings on krill stock hypotheses, with special attention to harvest rate and management procedures.Each observation will be explained in theoretical terms and, where possible, analyzed in detail. This approach reflects an ongoing work in progress that is continuously refined based on interactions with the CCAMLR Working Groups. These interactions provide valuable feedback that helps to shape and improve the study's methodology and findings. It should be noted, of this approach is vital for contributing to management strategies based on quantitative methods that provide indicators of krill health, integrating ecosystem aspects into the assessment.

\newpage

# METHODOLOGY

The study area for the assessment of krill populations in the Antarctic Peninsula is Subarea 48.1 and strata subdivision to account for spatial heterogeneity [@Dornam2021] (Figure \@ref(fig:mapa)). The Antarctic Peninsula region is vast and encompasses a variety of habitats and environmental conditions that can influence the distribution and abundance of krill. Therefore, dividing the study area into strata allows to compartmentalize information related to fishing and surveys to model krill populations by considering local variations in structure population.

## Data Sources

Multiple sources of data to inform the population dynamics model are included in this analysis. These data sources include: 1) Data from the monitoring program of the krill fishery were used, as CPUE and length compositions, which have been systematically collected on board fishing vessels by scientific observers as part of the CCAMLR Scheme of International Scientific Observation (SISO) program. This database has information about vessel, nationality, georeferenced, among others; 2) Data collected from scientific AMLR research cruises and monitoring programs specifically designed to assess krill populations, including: data on abundance by strata, and biological characteristics of krill collected through net sampling. All this data was processed in @Dornam2021; and length compositions from AMLR survey (provided by Ecosystem Science Division, NOAA) was handled according to the evaluation template requirements (Suppl. Mat. 2); 3) For this analysis, chlorophyll-a concentration (Chl) was assessed as a primary environmental driver. Chl data was sourced from Bio-Geo-Chemical, L4 (monthly and interpolated) satellite observations, obtained through the E.U. Copernicus Marine Service Information (doi.org/10.48670/moi-00021; doi.org/10.48670/moi-00148). This data provides a detailed understanding of primary productivity levels, which are critical for evaluating ecosystem health and  krill dynamics in the WAP. We fitted a GLMM to predict Length with Year (formula: Length ~ Year + Chl + Sea Ice Concentration (SIC) + Sea Surface Temperature (SST)). The model included spatial component (cell) as random effect (formula: ~1 | cell). Krill length compositions from fishery data was correlated with environmental variables by using Pearson correlations test, and General Linear Mixed Model (GLMM) [@Bates2015; @Makowski2020] (Suppl. Mat 3); 4) Predator abundance data from "MAPPPD". Although the penguin colonies that live within the study area behave differently in terms of dynamics throughout the analyzed series, we systematized a vectorized indicator for the purposes of considering it as a fleet to add like M2 in an integrated stock assessment.Data handling for modeling purposes was elaborated with *tidyverse* library  and dependences [@Wickham2019].

## Conceptual Model

A conceptual model for krill in SubArea 48.1 includes the five spatial strata described previously. Despite differences among strata, the krill population is closed for emigration and immigration, and their spatial distribution and interactions occur within this single, cohesive population (Figure \@ref(fig:concem)).Ecological and environmental variables that have an impact on the krill population include predator abundance (e.g., whales, seals, penguins), phytoplankton concentration, and nutrient levels. The model considers interannual fluctuations of growth dynamics , maturity and reproduction, with a spawning pulse per year.

## Statistical model Stock Synthesis (SS3)

Like most invertebrates, krill is a species for which aging is complex [@Punt2013]. Therefore, a size-structured model with age dynamics is used, which is transformed through a transition matrix based on the probabilities of age groups relative to the detailed structure. Integrated models have the capacity to reproduce the age dynamics of these populations and simultaneously transform them into population variables. The stock assessment model to krill was configured using Stock Synthesis (SS3 hereafter) [@Methot2013] with the updated version 3.30.21. SS3 is a widely used software tool for assessing fish and invertebrate populations widely used in the main regional fishing organizations to have approximations of the population dynamics of the exploited resource as well as to support management decisions. SS3 is a structured stock evaluation model, in the class of models called *"Integrated stock evaluation analysis model"* and has a set of sub-model that simulates growth, maturity, fecundity, recruitment,  and mortality processes, and observation, with expected values for different types of data. The model is coded in Automatic Differentiation Model Builder (ADMB) [@Fournier2012; @Methot2013: @methot2020stock] with estimation parameters. The methodology employed by Stock Synthesis involves a comprehensive and integrated approach, utilizing various data sources and modeling techniques to estimate the main population variables of krill in WAP, and unlike other integrated models, it has features that allow the integration of information related to the ecosystem, such as predators and environmental variables that have proven influence on the analyzed population. All analysis were executed in  R-CRAN [@R-base]and the graphical interface of the SS3 through *r4ss* [@Taylor2019]  and *ss3diags* packages [@Winker2023]. Life history parameters, like growth, weight-length relation, natural mortality and maturity was used as priors to model initial condition was taken from @Smith2023a, @Maschette2020 and @Kinzey2011 and can be found in Table \@ref(tab:parainit). Source data and temporal scale used in this stock assessment is showed in Figure \@ref(fig:datas2).

### Environmental variable modelling in `SS3`

One of the key challenges in this stock assessment framework is integrating ecosystem variables and drivers into the primary process of krill population dynamics and SS3 addresses these approaches [@methot2020stock]. The estimated level of krill recruitment depends on the spawning biomass from the previous season, an environmental time series—specifically, the 2000-2022 Chlorophyll-a time series—and a log-bias adjustment.

\[
E(\text{Recruitment}) = f(\text{SpBio}) \times \exp(B \times \text{envdata}) \times \exp\left(-0.5 \times \pi_R^2\right)
\]

\(R\) represents the variability of deviations, adding to the variance caused by environmental factors. \(SpBio\) represent spawning biomass, \(envdata\) is vector of time series of environmental variable. Consequently, as the environmental effect accounts for more of the total recruitment variability, the residual \(R\) should be reduced. However, the model does not automatically adjust for this.

Furthermore, this integrated model took into account environmental variables, as chlorophyll (Chl), known to influence krill abundance and reproductive outputs. These environmental factors play a critical role in shaping krill habitat suitability and productivity, thereby affecting their population dynamics. By integrating environmental data into our model, we aim to elucidate the relationships between environmental conditions and krill abundance, enabling more robust predictions of krill population dynamics under future climate scenarios.

### Predator components modelling in `SS3`

This stock assessment approach consider predator-prey interactions as a key driver of krill population dynamics. By incorporating information on predator  abundance and feeding ecology, our model provide a more comprehensive understanding of the tropic interactions shaping krill dynamics.  In this analysis we incorporated penguin population index as predator *"fleets"* as we describe previously. This overall natural morality (M) to include explicit mortality (M2) caused by each major predator as M = M1 + sum(M2) [@methot2020stock], and in this way transfer to total mortality in form Z = M1 + M2+ F.

All this component and path to construct models, diagnostics and check outputs is represented in Figure \@ref(fig:path).

\newpage

## Stock Assessment Scenarios

To consider the uncertainty associated with the krill nature scenarios, we considered a set of configurations associated with spatial heterogeneity, influence of life history parameters on the estimates and uncertainty associated with the relationship that exists between spawning biomass and krill recruitment. The methodology involves comparing the statistical performance of the model regarding these configurations. We have eight scenarios that were used to test different options in modeling about main considerations in assessment of krill population, and for comparative purposes, we consider `s1` as the base model (Table 1).

| Scenario | Description                                                                          |
|:------------:|:---------------------------------------------------------|
|   s01    | Fishery and Survey (AMLR) data, Predator, Environmental data, aggregated at 48.1 level               |
|    s1    | Fishery and Survey (AMLR) Length data, CPUE, Catch by strata. Predator and Env data |
|    s2    | "s1" without Stock-Recruit relation                                      |
|    s3    | "s1" Beverton & Holt S-R relation weak (0.9 steepness)                   |
|    s4    | "s1" BH S-R relation strong (0.6 steepness)   |
|    s5    | "s1" BH S-R relation mid-strong estimated    |
|    s6    | "s1" Ricker S-R relation estimated    |
|    s7    | "s1" w/ set of parameters estimated in @Mardones2024par   |

: Scenarios to modelling dynamics in krill

For the purposes of this working document, a comparative analysis of models `s01`, `s1` and `s7` will be evaluated, given that they contain elements such as aggregate data, spatially structured data and another set of life history parameters according to @Mardones2024par (unpublished).

\newpage

# RESULTS

## Diagnosis Base Model

All models tested was the final gradient using maximum likelihood is achieve and was relatively small (< 1.00e-04), and the Hessian matrix for the parameter estimates was positive, which shows that the performance is within what was expected for this type of analysis. About goodness-of-fit, Figure \@ref(fig:res1) and Figure \@ref(fig:res1a) display the residuals of length of krill captured by fleets over the years in `s1`. In the first graph, we observe multiple fishing fleets (FISHERYBS, FISHERYEI, FISHERYGS, FISHERYJOIN, FISHERYSSIW) and a survey (SURVEYBS). The second graph shows surveys (SURVEYEI, SURVEYGS, SURVEYJOIN) and predator data (PREDATOR) for `s1`. For many fleets and surveys, greater consistency in krill sizes is observed after the year 2000. However, there was variability in the width of the distributions over time, which may reflect changes in abundance or sampling selectivity. For instance, in FISHERYBS and FISHERYEI, the distributions widen and slightly shift towards larger sizes in recent years, indicating a possible change in the krill population or fishing practices. The residuals from the model fit suggest that the model captures the inter annual and inter-fleet lengths variability well, although some deviations indicate that additional unmodeled factors might be influencing the observed sizes (Figure \@ref(fig:resall),  \@ref(fig:fitcom)).

## Retrospective analysis

Retrospective analysis shows the pattern of bias that exists in the models and is one of the ways in which we have identified that in all the tested scenarios there is a pattern, however, model S1 is the one that performs better. @Carvalho2021b indicate that values of rho parameter that fall outside (-0.15 to 0.20) for SSB for longer-lived species, or outside (-0.22 to 0.30) for shorter-lived species indicates an undesirable retrospective pattern (Figure \@(fig:retros1)). To evaluate the overall model fit of the relative abundance indices and composition data in `s1`, we used the joint-index residual. Overall, `s1` had a good performance when evaluating the lengths compositions (14.2%), however  RMSE showed a low predictive power with respect to the indices (71.3%). A loess-smoother indicated there appeared to be increased variability in the residuals of model fit to CPUE over time (Figure \@ref(fig:rmses1)).

## Likelihood profile in spatial Model `s1`

For `s1` scenario, the gradient of the likelihood profile for the penalty on the index deviations was greater than other data sources. The second strongest gradient in the log-likelihood profile was observed for the length compositions Figure \@ref(fig:likepro). The gradient of the likelihood profile supported by the length-composition data is higher than those supported by the penalty for the recruitment deviates and CPUE indices.

Once the performance of the model was corroborated, it was possible to have an estimate and forecasting of the main population variables over the years in absolute terms, which can be seen in the Table \@ref(tab:mainvar)

\newpage

# RECOMMENDATIONS SAM - 23

The main recommendations were extracted from @CCAMLRWGSAM2024

*2.3 Noting that the authors recalled a previous independent review of an age-based integrated stock assessment for Antarctic krill which encouraged the development of such assessments (Thomson, 2016), the Working Group highlighted that the other reviewer on that panel noted that a length-based model could be considered due to the sparsity of direct age data (de Lestang, 2016). It also noted that this would avoid the approximations required when converting length data to age data. The Working Group further recalled that a similar comment had been made (WG-SAM-2023, paragraph 4.3) in relation to a pilot Casal2 age-based assessment (WG-SAM-2023/25). *

- **Response**:

We consider that a length-based model as an alternative to the age-based approach, especially given the scarcity of direct age data. We recognize that using a length-based model could reduce the need for approximations when converting length data to age data. We plan to explore this alternative by developing a length-based model to compare its performance and results with the current age-based model. Additionally, we will incorporate standard model diagnostics, as suggested, similar to those used in toothfish assessments, to ensure transparency and facilitate understanding of model performance.

Aging techniques for krill have not been available historically, although some recent progress has been made [@Kilada2017]. In this sense, when reading of hard parts is difficult like in krill, catch-at-length data are more plentiful, as the collection of length information is relatively cheap. Length offers insights into the age structure of the population, as there is a correlation between age and length. However, it's still necessary to convert catch-at-length data into catch-at-age data. Several methods are available to achieve this [@ICCA2003].

#### Age-Length Keys (ALKs)

Put simple, age-length keys (ALKs) are generated through the ageing of a sub-sample of the population, and used to convert larger length samples from a population into ages. ALKs describe distributions of size for each age, and the relative number of individuals at each age, i.e. they represent a matrix detailing the probability that a fish of a given length is of a particular age. Once such a key is available, samples of fish that were only measured for length can be distributed over age groups according to the key. The use of ALKs assumes that the sample of aged fish and the sample of fish measured for length are simple random samples from the same population. Then, the probability that a fish is of a particular age, given its length is the same for both samples.
The ALK should generally be applied to length data from the same time period, since variability in recruitment and survivorship at age will change the age-length composition over time, and hence the number of survivors at age used to weight the size at age compositions will vary. The ALK may need to be seasonal if growth is temporally distinctive or seasonal migration occurs. A single ALK should only be applied to size data from a number of years if growth is reasonably stationary, and the approach of Kimura and Chikuni (1987) is used. Suitable justification should be presented for multi-seasonal or multi-annual application of ALKs. It must also be noted that the application of an ALK derived from a single time of year can cause serious bias if used to compute catch at age for the entire year.
Age-at-length data from hard parts should be combined into suitable length groups for required gears, time periods and locations. The size of these length groups will depend on the spread of lengths found within the catch length frequencies, the growth rate of the species, and the variability in length-at-age. Length-stratified sampling is required to ensure that the required number of fish (determined by the variability of length-at-age) is available over all length groups. Developing ALKs represent is laborious, and hence the optimum collection of information is desirable. Formulae exist to estimate the number of age determinations and length measurements necessary to guarantee a given level of accuracy. Oeberst (2000) developed a universal cost function for ALKs, for example [@ICCA2003].

The proportion at age is calculated as:
$$
\frac{\text{Number at age for a length group}}{\text{Number of fish aged in that length group}}
$$

The ALK for the time period is raised to the length distribution for that time period:

$$
\text{Raised numbers at age by length group} = \text{Numbers at length} \times \text{Proportion at age for that length}
$$

If the ALK does not contain data for all the length groups in the length distribution then data in the length distribution may be assigned to adjacent length groups where data are present in the ALK. An appropriate age- length distribution by gear group and time period is then produced. Care must be taken since considerable biases can result, particularly at large lengths where individuals may be distributed over a wide range of ages.

The numbers at age by length group are summed over the length range to give numbers at age. The variances are also summed over the length groups and the two components labelled variance due to ageing and variance due to length sampling. This gives the age composition for the required time period. Numbers at age for all gears can be calculated as:

$$
\sum N_a \times \left(\frac{W_{ct}}{W_{cs}}\right)
$$
where $\sum N_a$ is the sum of sampled numbers at age, W~ct~ is the total commercial catch weight, and W~cs~ is the
sampled commercial catch weight.
Variance due to ageing of numbers at age for all gears can be calculated as:

$$
\sum \text{Var}_a
$$
where $\sum Var_a$  is the sum of variances due to ageing.
Variance due to length sampling of numbers at age can be calculated as:

$$
\sum \text{Var}_l
$$
where $\sum Var_a$  is the sum of variances due to length sampling.
The variances should be raised by:

$$
\frac{W_{ct}}{W_{cs}}
$$
where W~ct~ is the total commercial catch weight, and W~cs~ is the sampled commercial catch weight.

The proportions-at-age are then adjusted to find the best fit between the observed size frequency data and that predicted by the proportion-at-age and the ALK. The method can work well where the distributions of size for each age are close to those in the length frequency, but convergence can be slow. The reader is referred to the paper for more information.

Then Numbers-at-age in the model were converted to numbers-at-length (mm) using a von Bertalanffy growth relationship withestimated parameters L~inf~, k, and v(Eq. (A.6)). This length-at-age relationship, mediated by survey catchability and selectivity (Eqs.(A.2) and (A.3)), was used to compare survey biomasses with model estimates of vulnerable biomass (Eq. (A.4)). Weight-at-length was assumed known and calculated using Eq. 3 from Hewitt et al. (2004)(W(L) = 2.236 × 10-6L3.314). (REF Kindzey)

Another way that we implemented in `SS3` is regarding *platoons*. This approach allows for the exploration of survivorship that varies with size. A value of 1 will not create additional groups. Odd-numbered values (e.g., 3, 5) will divide the overall morph into that many groups, resulting in smaller, larger, and average growth groups. Increasing the number of groups slows down the model's performance, so it's recommended not to exceed a value of 5. The fraction of each morph assigned to each group can be either user-defined or approximated using a normal distribution. When multiple groups are designated, an additional input is required to specify the ratio of variability in size-at-age between groups versus within groups. This ratio is used to distribute the total growth variability. The size-at-age for each group is then calculated as a factor (based on the between-group and within-group variability) times the size-at-age of the central morph, which is determined from the growth parameters specific to that Growth Pattern x Sex combination [@methot2020stock].

The age specific rates of fishing mortality can be linked to length specific selectivity. If growth platoons are invoked, then each platoon can have different age specific F due to interaction of length selectivity with their different size at age (Methot, *com pers*)

In `SS3` colud be conditioned in this way;

```{r eval=FALSE, echo=TRUE}
#_N_Growth_Patterns (Growth Patterns, Morphs, Bio Patterns, GP are terms used interchangeably in SS)
1  
#_N_platoons_Within_GrowthPattern
3 
 #_Platoon_between/within_stdev_ratio (no read if N_platoons=1)
0.4
#vector_platoon_dist_(-1_in_first_val_gives_normal_approx)
c(0.237, 0.464, 0.237)
```

```{r platoon, out.width='70%', fig.cap="Reprsentation of platoons in Krill lenght frencuecies"}
knitr::include_graphics('Figs/Plattoons.png')
```

Regarding to observation in SAM -2023 (paragraph 4.3) that say;

*4.3 its usefulness in providing an additional approach to assessing krill stock status, it recalled that the development of an integrated stock assessment for krill was considered desirable within three to five years"* but considering this;

*(i) the impact of using an age-based or a length-based implementation could be assessed, noting that a preservation of the source data would be preferred (i.e. the conversion between length composition and age composition should be considered, with conversions from length to age to length avoided)*

*It encouraged the authors to provide standard model diagnostics, similar to those presented for toothfish assessments, to facilitate understanding of model performance (e.g., WG-SAM-2023, paragraphs 6.33 and 6.34).*

*2.4 While noting that this study constituted interesting and important work, the Working Group noted that some issues needed further consideration including the change of trawl designs over the course of the collection of the data used as inputs to the model (e.g. changes in mesh sizes and the presence of fine-mesh codend inserts) and the likely invalid assumption of the Peninsula as being a closed system. It further highlighted the need to discuss the development of a standardized data collection plan to support the ongoing revision of the krill fishery management approach.*

  - **Response**:

We appreciate the identification of potential issues related to changes in trawl designs and the assumption of the Peninsula as a closed system. We will review the trawl data to assess the impact of changes in mesh sizes and fine-mesh codend inserts on the model inputs. Moreover, we acknowledge the need to reconsider the assumption of the Peninsula being a closed system, and we will explore alternative approaches to address this. Moving forward, we will engage in discussions to develop a standardized data collection plan that aligns with the revision of the krill fishery management approach, ensuring that future data inputs are consistent and reliable.

*2.5 Dr S. Kasatkina (Russian Federation) noted that data on krill length and biological composition from catches of fishing vessels will be not suitable for such a modelling approach, and recalled that comparisons of krill length composition from catches of fishing trawls and catches of scientific trawls within the same fishing ground revealed significant differences (WG-ASAM-2021/03). Moreover, Dr Kasatkina noted that there were significant differences in the length composition of catches between fishing vessels and these differences are random in nature, which may be due to both the selective properties of commercial trawls, different fishing methods (continuous and traditional fishing) as well as the efficiency of krill sampling by observers at-sea (WG-ASAM-2021/03; WG-EMM-2024/37). Dr Kasatkina noted that this modelling effort required clarity regarding the interaction between the fishery and krill-dependent predators, which requires regular observations to study the spatial overlap of fishing zones and predator foraging zones and could be accompanied by krill distribution patterns (for example, such complex ecosystem observations were provided on RV Atlantida in 2020; SC-CAMLR-42/07). Dr Kasatkina recalled that the spatial and temporal dynamics of groups from the Bellingshausen and Weddell Seas (Fach et al, 2002; Murphy et al, 2004; WG-EMM-2024/43; WG-EMM-2024/39) and noted that without data on krill transport and standardised acoustic surveys of krill, it is impossible to assess the influence of factors such as spatial heterogeneity and life history parameters on key krill population variables.*

- **Response**: We understand the concerns raised regarding the suitability of krill length and biological composition data from fishing vessels for this modeling approach. We agree that significant differences in length composition between commercial and scientific trawls, as well as among different vessels, pose challenges. To address this, we will review the selection criteria for the data used in our model and consider incorporating data exclusively from standardized scientific trawls to improve the reliability of the model outputs. We also acknowledge the importance of understanding the interaction between fisheries and krill-dependent predators. We will explore opportunities to integrate regular observations and data on the spatial overlap of fishing zones and predator foraging areas into the model, possibly leveraging existing ecosystem observations like those from RV Atlantida.

*2.6 The Working Group encouraged the authors to provide a progression of model implementations from simple to more complex to facilitate understanding and evaluate the evidence for the model assumptions. It further noted the presence of patterns in the residuals shown in the paper (Figures 5 and 6) which warranted further investigation, as well as the need to assess the realism of some parameter values used in the model. The Working Group also indicated that this work would benefit from taking into consideration recent findings by SKEG regarding the krill stock hypothesis (e.g. WG-EMM-2024/39).*

 - **Response**: We agree that a progression from simple to more complex model implementations would facilitate understanding and provide a clearer evaluation of the model assumptions. We will outline a plan to present this progression in future work, beginning with a simpler model and gradually incorporating complexity. Furthermore, we acknowledge the presence of patterns in the residuals and the need to investigate these further. We will conduct a thorough analysis of the residuals to identify any underlying issues and reassess the realism of the parameter values used in the model. In addition, we will incorporate recent findings from SKEG regarding the krill stock hypothesis to ensure that our model assumptions are up-to-date and reflect the latest scientific understanding.

*show  Harvest Rate*

*Diagram management procedure*

\newpage

# DISCUSSION

The analysis presented here represents a modern approach for stock assessment of complex dynamics such as krill and the integration of ecosystem components as environmental variables and ecological aspects such as predators. The development of an integrated stock assessment model like could provide support to adaptive management, wherein management procedures resulting, like status or quota, are adjusted based on ecosystem monitoring indices [@Wang2021]. However, integrated models are not new to krill in SO. For several years, CCAMLR has been developing and applying integrated assessment methodologies that incorporate diverse data sources, such as acoustic and trawl surveys, to improve the accuracy of their stock assessments. In 2016, through an external review process,@Thomson2016 explicitly recommended *"the use of an Integrated Analysis model for assessment of Antarctic krill in FAO Subarea 48.1 is appropriate and would constitute an improvement over the modeling strategies currently in use to manage krill"* and he stated that the an integrated modeling framework *"has the potential to substantially improve the scientific basis for future krill management"*.

@Kinzey2015a and @Kinzey2019a used integrated models with survey data to test the impacts of selectivity and recruitment estimation in certain areas of Subarea 48.1. @Kinzey2011 implemented a model to test spawning biomass, recruitment and movement between areas associated with the northern Antarctic Peninsula. Although these models had convergence issues, they represented significant advances in this type of framework. Additionally, @Wang2021 made contributions to integrated statistical models with survey and fishery data for the northern sector of the Western Antarctic Peninsula, focusing on the Bransfield Strait, testing different data weightings. Furthermore, examined correlations between recruitment and climatic variables such as the Southern Annular Mode (SAM) El Niño Oscillation Southern (ENSO). In absolute values and with respect to biomass levels, although the estimates of our model were lower than the previously described models, there is a coincidence with respect to recruitment pulses as well as mortality levels.

Spatial heterogeneity is a fundamental aspect of krill ecology, driven by a myriad of factors including oceanographic conditions and predator foraging behavior. By incorporating spatial heterogeneity into our model considering the strata as part of the analysis, we aim to capture the complex spatial patterns of krill distribution. This spatially implicit approach will allow us to identify krill hotspots, areas of high productivity, and potential ecological corridors, providing valuable insights for conservation planning and management strategies [@Watters2013] and transfer those signals of differences in krill population structure into the model to generate its estimates.

From an ecological point of view, the study area encompasses the primary spawning, breeding, and wintering regions of Antarctic krill in the southwest Atlantic Ocean. There are many studies that confirm the influence of predation on krill populations within the Antarctic Peninsula [@Silvestro2023; @Reisinger2022]. For this, we consider the penguin  population as one of the most powerful elements of natural predation on krill because they have direct correlations between this type of taxonomic group and krill  [@Kruger2021]. The intention to incorporate elements of predation in a stock is necessary to improve the modeling, because it did not account for uncertainty in natural mortality underestimated uncertainty in current stock biomass by as much as 20% [@Hollowed2000].

In a scientifically rigorous stock assessment, this analysis included comparisons between various natural scenarios and assumptions used to model krill population dynamics. It is important to note that the models presented here can be refined in terms of initial conditioning to improve fits on length compositions of predators and indices that were initially low. However, a primary objective was to evaluate the impact on key population variables, considering factors such as spatial heterogeneity and life history parameters. This impact was verified in this stock assessment approach, identifying aspects that must be considered when making recommendations for krill management by CCAMLR using models.

By comparing model outputs under scenarios where spatial heterogeneity is assumed versus not assumed, the significance of incorporating spatial variation into the model can be assessed. This methodology enables researchers to make informed decisions about the necessity of accounting for spatial heterogeneity in understanding krill distribution and abundance, thereby enhancing the accuracy and reliability of ecological modeling efforts.

Nowadays, one of the frontiers of stock assessment is the incorporation of ecosystem and spatial considerations into modeling. In this context, the features of the SS3 platform allow us to advance in this direction and contribute to the discussion, both in understanding the population dynamics of krill and providing advice for sustainable management. In summary, our proposed stock assessment scheme represents a pivotal approach to understanding the complex population dynamics of Antarctic krill in Subarea 48.1.

\newpage

# SUPPLEMENTARY MATERIAL

- Supplementary Material 1:  Model template example and initial condition to `s1` scenario can be found in this [repo](https://github.com/MauroMardones/SA_Krill/tree/main/test) and the executable version (3.30.21) can be download from NOAA Virtual Lab [SS3](https://vlab.noaa.gov/web/stock-synthesis)

- Supplementary Material 2:  Length composition templates preparation to SS3 from AMLR can be found [link](https://mauromardones.github.io/AMLR_Length_Data/) and routine code in this [repo](https://github.com/MauroMardones/AMLR_Length_Data)

- Supplementary Material 3:  Correlation analysis between krill biology and environmental variables can be found in this [link](https://mauromardones.github.io/Krill_Length_Cor/) and routine code in this [repo](https://github.com/MauroMardones/Krill_Length_Cor)

\newpage

# FIGURES AND TABLES

```{r mapa, out.width='40%', fig.show='hold',fig.cap="Subarea 48.1 and management strata considered in the spatio-temporal analysis of intrinsic productivity of Krill (BS=Brainsfield Strait, EI= Elephant Island, Gerlache= Gerlache strait, JOIN= Joinville Island, SSWI= South West)"}
knitr::include_graphics('Figs/481.png')
```

```{r concem, out.width='40%', fig.cap="Conceptual model used to model dynamics population in Antarctic krill in WAP"}
knitr::include_graphics('Figs/conceptual.jpeg')
```

```{r datas2, out.width='70%', fig.cap="Source information time series for each fleet used in s1 scenario"}
knitr::include_graphics('Figs/data_plot_s2.png')
```

```{r parainit, fig.cap="Initial biology and fishery parameters to set a model s1 in krill in WAP"}

# leo archivos para plotear y hacer tablas
start1 <- SS_readstarter(file = file.path(dir1,
                                          "starter.ss"),
                              verbose = FALSE)
# note the data and control file names can vary, so are determined from the 
# starter file.
dat1 <- SS_readdat(file = file.path(dir1, start1$datfile),
                        verbose = FALSE)
# Read in ctl file. Note that the data fileR object is needed so that SS_readctl
# assumes the correct data structure
ctl1 <-  r4ss::SS_readctl(file = file.path(dir1,
                                    start1$ctlfil),
                        verbose = FALSE,
                        use_datlist = TRUE, 
                   datlist = dat1)
fore1 <- r4ss::SS_readforecast(file = file.path(dir1, 
                                                "forecast.ss"),
                              verbose = FALSE)
# can also read in wtatage.ss for an empirical wt at age model using
# r4ss::SS_readwtatage()

parbio<-ctl1$MG_parms[1:10,c(1:4,7)]
 row.names( parbio)<-c("Nat M",
                       "Lmin", 
                       "Lmax",
                       "VonBert K",
                       "CV young",
                       "CV old", 
                       "Wt a", 
                       "Wt b",
                       "L50%", 
                       "Mat slope")

 SRpar<-ctl1$SR_parms[1:5,c(1:4,7)]
 Qpar<-ctl1$Q_parms[1:2,c(1:4,7)]
 Selpar<-ctl1$size_selex_parms[1:22,c(1:4,7)]
 parInit<-as.data.frame(rbind(parbio,SRpar,Qpar,Selpar))

parInit %>%
  kbl(booktabs = TRUE,
      format = "latex",
      position="ht!",
    caption = "Input parameters for the initial SS3 model of krill. Each parameter line contains a minimum value (LO), maximum value (HI), and initial value (INIT). If the phase (PHASE) for the parameter is negative, the parameter is fixed as input") %>%
  kable_paper("hover", 
              full_width = F)%>%
  kable_styling(latex_options = c("striped"),
                full_width = FALSE,
                font_size=9)%>% 
  pack_rows(index = c("Natural Mortality" = 1,
                        "Growth"= 5,
                        "Length-Weigth Relation" = 2,
                        "Maturity"=2,
                        "Stock-Recruit Relation"=5,
                        "Catchability"=2,
                        "Selectivity"=4))
```

\begin{landscape}

```{r path, fig.cap="Framework path to stock assessment model in krill in WAP (Yellow boxes is not implemeted yet)."}
knitr::include_graphics('Figs/pathmod.png')
```

\end{landscape}

```{r res1, fig.cap = "Pearson residuals, comparing across fleets. Closed bubbles are positive residuals (observed > expected) and open bubbles are negative residuals (observed < expected)."}
knitr::include_graphics('s1/plots/comp_lenfit__page1_multi-fleet_comparison.png')
```

```{r res1a, fig.cap = "Pearson residuals, comparing across fleets. Closed bubbles are positive residuals (observed > expected) and open bubbles are negative residuals (observed < expected) (continued)."}
knitr::include_graphics('s1/plots/comp_lenfit__page2_multi-fleet_comparison.png')
```

```{r resall, out.width='30%', fig.show='hold', fig.cap="Mean length for each with 95% confidence intervals based on current sample sizes. blue line represent estimated"}
par(mfrow=c(5,2))
knitr::include_graphics(c('s1/plots/comp_lenfit_data_weighting_TA1.8_FISHERYBS.png',
                          's1/plots/comp_lenfit_data_weighting_TA1.8_FISHERYEI.png',
                          's1/plots/comp_lenfit_data_weighting_TA1.8_FISHERYGS.png',
                          's1/plots/comp_lenfit_data_weighting_TA1.8_FISHERYJOIN.png',
                          's1/plots/comp_lenfit_data_weighting_TA1.8_FISHERYSSIW.png',
                          's1/plots/comp_lenfit_data_weighting_TA1.8_PREDATOR.png',
                          's1/plots/comp_lenfit_data_weighting_TA1.8_SURVEYBS.png',
                          's1/plots/comp_lenfit_data_weighting_TA1.8_SURVEYEI.png',
                          's1/plots/comp_lenfit_data_weighting_TA1.8_SURVEYGS.png',
                          's1/plots/comp_lenfit_data_weighting_TA1.8_SURVEYJOIN.png'))

```

```{r fitcom, out.width='80%', fig.show='hold', fig.cap="Fits to lengths compositions in different fleets in s1 scenario"}
knitr::include_graphics('s1/plots/comp_lenfit__aggregated_across_time.png')

```

```{r retros1, out.width='70%', fig.cap="Retrospective pattern in s1 scenario in Antarctic krill"}
knitr::include_graphics('Figs/retros1.png')
```

```{r rmses1, out.width='60%', fig.show='hold',  fig.cap="Joint residual plots for index (upper panel) and lengths (bottom panel) fits from different fleets in Antarctic krill"}
par(mfrow=c(2,1))
knitr::include_graphics(c('Figs/rmse21.png',
                          'Figs/rmselens1.png'))
```

```{r likepro, out.width='70%', fig.cap="Log-likelihood profiles for R0 for the various data components included in Antarctic krill"}
knitr::include_graphics('Figs/Profile_s1.png')
```

```{r mainvar, fig.cap="Main variables outputs (Total Biomass, Biomass at 75%, Spawining biomass and Recruit in t) from estimation in s1 model krill in WAP"}
out_s2 <- read_excel("DataKrill.xlsx", 
    sheet = "variable_s2")
out_s2 %>%
  kbl(booktabs = TRUE,
      format = "latex",
    caption = "Main variables outputs from stock asssessment krill in WAP") %>%
  kable_paper("hover", 
              full_width = TRUE)%>%
  kable_styling(latex_options = c("striped",
                                  "condensed"),
                full_width = TRUE,
                font_size=6 )#%>% 
  #pack_rows(index = c("Estimation" = 1,
   #                     "Prediction" = 45))
```

\newpage
```{r recdev, out.width='90%', fig.cap="Recruit deviation for all scenarios in krill population"}
knitr::include_graphics('index_files/figure-html/unnamed-chunk-47-1.jpeg')
```

```{r acf, out.width='90%', fig.cap="Cross correlation in autoregresive analysis for all scenarios in krill population"}
knitr::include_graphics('index_files/figure-html/unnamed-chunk-46-1.jpeg')
```

```{r likeall, out.width='90%', fig.cap="Components of total likelihood for all sources and all tested scenarios"}
knitr::include_graphics('Figs/Likelihoodtotal.png')
```

```{r longterm, out.width='100%', fig.cap="Long term estimation for different scenarios in krill spawning biomass for whole time series (1979-2020"}
knitr::include_graphics('Figs/longtermsb.png')
```

\newpage

\begin{landscape}

```{r comparision, out.width='40%', fig.show='hold', fig.cap="Comparative perfomance with scenarios s01, s1 and s7 to recruit, spawing biomass, biomass ratio (hopefully equal to fraction of unfished) and Virgin biomas probability"}
par(mfrow=c(2,2))
knitr::include_graphics(c('Figs/compare10_recruits_uncertainty.png',
                          'Figs/compare2_spawnbio_uncertainty.png',
                          'Figs/compare4_Bratio_uncertainty.png',
                          'Figs/compare17_densities_SSB_Virgin.png'))
```

\end{landscape}

\newpage

# REFERENCES