### 2025-03-27 index.Rmd
adding stest to MSE

---
title: "Supplementary Information 1"
subtitle: "Assessing ecosystem effects on the stock productivity of Antarctic Krill (*Euphausia superba*): An Integrated modeling perspective"
author: "Mardones, M; Jarvis Mason, E.T.; Pinones, A.;  Santa Cruz, F.; Cárdenas, C.A"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
bibliography: SA_krill.bib
#csl: apa.csl
csl: icesjournal.csl
link-citations: yes
linkcolor: blue
output:
  html_document:
    keep_md: true
    toc: true
    toc_deep: 3
    toc_float:
      collapsed: false
      smooth_scroll: false
    theme: cosmo
    fontsize: 0.9em
    linestretch: 1.7
    html-math-method: katex
    self-contained: true
    code-tools: true
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup1, echo=FALSE}
#rm(list = ls())
#set.seed(999)
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      fig.align = 'center',
                      fig.pos = "H",
                      dev = 'jpeg',
                      fig.path = "Figs/",
                      dpi = 300,
                      tidy.opts=list(width.cutoff=50), 
                      tidy=TRUE)
#XQuartz is a mess, put this in your onload to default to cairo instead
options(bitmapType = "cairo") 
# (https://github.com/tidyverse/ggplot2/issues/2655)
# Lo mapas se hacen mas rapido
```

```{r message=FALSE, eval=TRUE}
# install.packages("devtools")
# devtools::install_github("r4ss/r4ss", ref="development")
# install.packages("caTools")
# library("caTools")
# install.packages("r4ss")
library(r4ss)
library(here)
#remotes::install_github("PIFSCstockassessments/ss3diags")
library(ss3diags)
library(kableExtra)
library(doParallel) # facilita la ejecución paralela en R
detectCores()
registerDoParallel(8)
library(ggpubr)
library(tibble)
library(openxlsx)
library(ggthemes)
library(forecast)
library(tidyr)
library(mixR)
library(readxl)
library(tidyverse)
library(ggridges)
```

```{r message=FALSE, warning=FALSE, echo=TRUE}
# dir01 <- here("s01") # agreggate data (no spatial diferences)
# dir1<-here("s1") # Data strata flishery
# dir2<-here("s2") # Same 9 with areas (SubStrata) as fleet. Dif size comoposition and dif CPUE and dif survey length and biomass data by strata
# dir3<-here("s3") # without S-R
# dir4<-here("s4") # 
# dir5<-here("s5") # 
# dir6<-here("s6") # 
# dir7<-here("s7") # 2 set parametres EMM-2024/23 (Mardones)
# dir8<-here("s8") # s1 platoons
# dir9<-here("s9") # s1 w/ blocks
dir1.1<-here("s1.1") # con todo
dir1.2 <- here("s1.2") # sin predador ni ambiental
dir1.3 <- here("s1.3") # borrado los registros de prreddador )igual q el S2
Figs <- here("Figs")# S
```

## Overview

This study aims to evaluate the impact of ecosystem components—such as environmental variables and predator-prey interactions—on the productivity and key population dynamics of *Euphausia superba* in Subarea 48.1. By incorporating these factors into the assessment, we analyze how krill population variables respond to ecological variability, providing insights into their resilience and potential management implications.

Here, the **reference model** represents a baseline assessment of *Euphausia superba* population dynamics in Subarea 48.1, excluding environmental and ecological variables. This model assumes that krill productivity and population parameters are driven  by intrinsic biological processes, such as growth, mortality, and recruitment and fishery impacts without accounting for external influences like environmental variability or predation pressure. By serving as a *control scenario*, this model provides a benchmark against which the impact of ecosystem components in productivity can be evaluated, allowing for a direct comparison of how environmental and ecological factors influence krill stock dynamics.

## Statistical Model (SS3)

Stock Synthesis (v.3.30.21)  is a widely used tool for assessing fish and invertebrate populations, including Antarctic krill. SS3 is implemented in `C++` with estimation enabled through automatic differentiation (ADMB) [@Fournier2012; @Methot2013]. In this exercise, SS3 is configured as an integrated stock assessment model, explicitly accounting for age and size structure while incorporating key ecosystem drivers. The model simulates population processes such as growth, maturity, fecundity, recruitment, movement, and mortality, while also integrating environmental variability and predator-prey relationships to refine estimates of population trends. The analysis of model outputs is conducted using R, utilizing the *r4ss* and *ss3diags* packages [@Taylor2019; @Winker2023]. By leveraging a spatially implicit, ecosystem-informed approach, this assessment provides a robust framework for evaluating krill stock dynamics under changing environmental conditions. These insights are crucial for informing sustainable management strategies in the Antarctic Peninsula region, where krill plays a foundational role in the marine food web.

## Parameters

The following table summarizes the key parameters to conditioning the reference model, including biological, growth, and population dynamics factors.

```{r}
start1 <- SS_readstarter(file = file.path(dir1.1,
                                          "starter.ss"),
                              verbose = FALSE)
dat1 <- SS_readdat(file = file.path(dir1.2, start1$datfile),
                        verbose = FALSE)
ctl1 <-  r4ss::SS_readctl(file = file.path(dir1.1,
                                    start1$ctlfil),
                        verbose = FALSE,
                        use_datlist = TRUE, 
                   datlist = dat1)
fore1 <- r4ss::SS_readforecast(file = file.path(dir1.1, 
                                                "forecast.ss"),
                              verbose = FALSE)
# can also read in wtatage.ss for an empirical wt at age model using
# r4ss::SS_readwtatage()
```

```{r}
parbio<-ctl1$MG_parms[1:10,c(1:3,7)]
 row.names( parbio)<-c("Nat M",
                       "Lmin", 
                       "Lmax",
                       "VonBert K",
                       "CV young",
                       "CV old", 
                       "Wt a", 
                       "Wt b",
                       "L50%", 
                       "Mat slope")

 SRpar<-ctl1$SR_parms[1:5,c(1:3,7)]
 Qpar<-ctl1$Q_parms[1:2,c(1:3,7)]
 Selpar<-ctl1$size_selex_parms[1:22,c(1:3,7)]
 parInit<-as.data.frame(rbind(parbio,SRpar,Qpar,Selpar))

wb <- createWorkbook()

addWorksheet(wb, "parameters")
writeData(wb, "parameters", parInit)

# Guardar el workbook
#saveWorkbook(wb, "DataKrill.xlsx", overwrite = TRUE)

```

```{r}
parInit %>%
  mutate(across(where(is.numeric), ~ round(., 3)))  %>% 
  kbl(booktabs = TRUE,
      format = "html",
      position = "ht!",
      caption = "Input parameters for the initial SS3 model of krill. Each parameter line contains a minimum value (LO), maximum value (HI), and initial value (INIT). If the phase (PHASE) for the parameter is negative, the parameter is fixed as input") %>%
  kable_paper("hover", full_width = TRUE) %>%  # Expande el ancho
  kable_styling(latex_options = c("striped", "condensed"),
                full_width = TRUE,  # Expande la tabla al ancho completo
                font_size = 12,
                html_font = "Times New Roman") %>%  # Aplica la fuente Times New Roman
  pack_rows(index = c("Natural Mortality" = 1,
                      "Growth" = 5,
                      "relationship Length-Weigth" = 2,
                      "Maturity" = 2,
                      "S-R relation" = 5,
                      "Catchability" = 2,
                      "Selectivity" = 4))

```
Source of data inpit

```{r fig.width=9, fig.height=4}
dat1_std <- dat1$CPUE %>%
  group_by(index) %>%
  mutate(index_std = obs / mean(obs, na.rm = TRUE)) %>%
  ungroup() %>% 
  mutate(index = recode(index,
                        `1` = "FISHERYBS",
                        `2` = "FISHERYEI",
                        `3` = "FISHERYGS",
                        `4` = "FISHERYJOIN",
                        `5` = "FISHERYSSIW",
                        `6` = "SURVEYBS",
                        `7` = "SURVEYEI",
                        `8` = "SURVEYGS",
                        `9` = "SURVEYJOIN",
                        `10` = "SURVEYSSIW",
                        `11` = "PREDATOR"))
ggplot(dat1_std, aes(x = year, y = index_std, group=index))+
  geom_point() +
  geom_smooth(method="lm",
              se=F)+
  facet_wrap(.~index,
             scales = "free_y",
             ncol=6)+
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, 
                                   hjust = 1))+
  labs(title = "",
       x = "",
       y = "Standar index",
       color = "index")
```

```{r fig.width=10, fig.height=4}
dat_long <- dat1$lencomp %>% 
  pivot_longer(cols = starts_with("l"),  
               names_to = "Talla",  
               values_to = "Frecuencia") %>% 
  mutate(Talla = as.numeric(sub("l", "", Talla))) %>%  
  mutate(FltSvy = recode(FltSvy,
                         `1` = "FISHERYBS",
                         `2` = "FISHERYEI",
                         `3` = "FISHERYGS",
                         `4` = "FISHERYJOIN",
                         `5` = "FISHERYSSIW",
                         `6` = "SURVEYBS",
                         `7` = "SURVEYEI",
                         `8` = "SURVEYGS",
                         `9` = "SURVEYJOIN",
                         `10` = "SURVEYSSIW",
                         `11` = "PREDATOR")) %>%
  mutate(FltSvy = factor(FltSvy, levels = c("FISHERYBS", 
                                            "FISHERYEI", 
                                            "FISHERYGS",
                                            "FISHERYJOIN", 
                                            "FISHERYSSIW",
                                            "SURVEYBS",
                                            "SURVEYEI", 
                                            "SURVEYGS", 
                                            "SURVEYJOIN",
                                            "SURVEYSSIW",
                                            "PREDATOR")))

mean_talla <- dat_long %>%
  group_by(FltSvy) %>%
  summarise(mean_Talla = mean(Talla, na.rm = TRUE))

# Expandir datos según la frecuencia
dat_expanded <- dat_long %>%
  uncount(weights = Frecuencia)

# Crear el gráfico con líneas verticales en cada facet
ggplot(dat_expanded, aes(x = Talla, y = as.factor(Yr), 
                         fill = stat(x))) +
  geom_density_ridges_gradient(scale = 1.8,
                               rel_min_height = 0.001,
                               col = "black",
                               fill = "grey") +
  facet_wrap(~FltSvy, ncol = 10) +
  geom_vline(data = mean_talla, aes(xintercept = mean_Talla), 
             color = "red", 
             size = 0.7,
             alpha=0.5) +
  labs(title = "",
       x = "cm",
       y = "",
       fill = "Densidad") +
  theme_few()
```

## Scenarios

In Table 1 we have ten scenarios to test different option in modeling
about main consideration in assessment of krill population.

| Scenario | Description                                                                          |
|:------------:|:---------------------------------------------------------|
|    s1.1  | Spatial data without environmental and predator components                           |
|    s1.2  | "s1.1" with environmental and predator components                                    |
|    s1.3  | "s1.2" w/ selectivity blocks                                                         |

## Run Models

```{r eval=FALSE, message=F, include=FALSE, echo=TRUE}
directorios <- c("s1.1",
                 "s1.2",
                 "s1.3")

for (dir in directorios) {
  r4ss::run(
    dir = dir,
    exe = "ss_osx",
    skipfinished = FALSE,
    show_in_console = TRUE
  )
}
```

```{r eval=FALSE, message=F, include=FALSE}
#OR with rby separate
#shell(cmd="ss") # run SS to windows
# or 
r4ss::run(
  dir = dir1.1,
  exe = "ss_osx",
  skipfinished = FALSE, # TRUE will skip running if Report.sso present
  show_in_console = TRUE # change to true to watch the output go past
)

# s1.2
r4ss::run(
  dir = dir1.2,
  exe = "ss_osx",
  skipfinished = FALSE, # TRUE will skip running if Report.sso present
  show_in_console = TRUE # change to true to watch the output go past
)

# s1.3
r4ss::run(
  dir = dir1.3,
  exe = "ss_osx",
  skipfinished = FALSE, # TRUE will skip running if Report.sso present
  show_in_console = TRUE # change to true to watch the output go past
)

```

```{r message=F, include=FALSE}
#s1.1
base.model1.1 <- SS_output(dir=dir1.1,
                        covar=T,
                        forecast=T)
#s1.2
base.model1.2 <- SS_output(dir=dir1.2,
                        covar=T,
                        forecast=T)
#s1.3
base.model1.3 <- SS_output(dir=dir1.3,
                        covar=T,
                        forecast=T)
```

Data used en both (spatial and No spatial models)

```{r }
SSplotData(base.model1.1, 
           subplot = 1,
           pheight = 15)

SSplotData(base.model1.2, 
           subplot = 1,
           pheight = 15)
```
# RESULTS

```{r eval=FALSE, message=F, include=FALSE}
# Definir la ruta de la carpeta
folder_path <- "s1.1/plots"
# Verificar si la carpeta existe y eliminarla
if (dir.exists(folder_path)) {
  unlink(folder_path, recursive = TRUE)
  message("La carpeta 'plots' ha sido eliminada.")
} else {
  message("La carpeta 'plots' no existe.")
}

SS_plots(base.model1.1, 
         uncertainty = TRUE,
         datplot = T, 
         png=T, 
         aalresids = F,
         btarg=0.75, 
         minbthresh=0.2, 
         forecast=T)

# S 1.2

# Definir la ruta de la carpeta
folder_path2 <- "s1.2/plots"
# Verificar si la carpeta existe y eliminarla
if (dir.exists(folder_path2)) {
  unlink(folder_path2, 
         recursive = TRUE)
  message("La carpeta 'plots' ha sido eliminada.")
} else {
  message("La carpeta 'plots' no existe.")
}

SS_plots(base.model1.2, 
         uncertainty = TRUE,
         datplot = T, 
         png=T, 
         aalresids = F,
         btarg=0.75, 
         minbthresh=0.2, 
         forecast=T)

# S 1.3

# Definir la ruta de la carpeta
folder_path3 <- "s1.3/plots"
# Verificar si la carpeta existe y eliminarla
if (dir.exists(folder_path3)) {
  unlink(folder_path3, 
         recursive = TRUE)
  message("La carpeta 'plots' ha sido eliminada.")
} else {
  message("La carpeta 'plots' no existe.")
}

SS_plots(base.model1.3, 
         uncertainty = TRUE,
         datplot = T, 
         png=T, 
         aalresids = F,
         btarg=0.75, 
         minbthresh=0.2, 
         forecast=T)
```

### Main Variables poulation in `s1.1` scenario

```{r}
SSplotTimeseries(base.model1.1,
                 subplot = 1)
SSplotTimeseries(base.model1.1,
                 subplot = 9)
SSplotTimeseries(base.model1.1,
                 subplot = 7)
```

### Main Variables poulation in `s1.2` scenario

```{r}
SSplotTimeseries(base.model1.2,
                 subplot = 1)
SSplotTimeseries(base.model1.2,
                 subplot = 9)
SSplotTimeseries(base.model1.2,
                 subplot = 7)
```

### Main Variables poulation in `s1.3` scenario

```{r}
SSplotTimeseries(base.model1.3,
                 subplot = 1)
SSplotTimeseries(base.model1.3,
                 subplot = 9)
SSplotTimeseries(base.model1.3,
                 subplot = 7)
```

Selectivity

```{r fig.height=6}
SSplotSelex(base.model1.1,
            subplots = 1)

SSplotSelex(base.model1.2,
            subplots = 1)
```

```{r}
SSplotBiology(base.model1.1,
            subplots = 1:32)
```

```{r}
SSplotComps(base.model1.2,
            subplots = 1)

```

Total biomass

```{r eval=FALSE}
total01 <- as.data.frame(base.model1.1$timeseries[1:31,c(2,5)]) %>%
  mutate(Serie = "No Spatial")
total1 <- as.data.frame(base.model1.2$timeseries[,c(2,5)]) %>% 
  mutate(Serie = "Spatial implicit")
total2 <- as.data.frame(base.model2$timeseries[,c(2,5)]) %>% 
  mutate(Serie = "Spatial w/ without S-R relation")
total3 <- as.data.frame(base.model3$timeseries[,c(2,5)]) %>% 
  mutate(Serie = "Spatial w/ S-R relation weak (0.9 steepness)")
total4 <- as.data.frame(base.model4$timeseries[,c(2,5)]) %>% 
  mutate(Serie = "Spatial w/ S-R relation strong (0.6 steepness)")
total5 <- as.data.frame(base.model5$timeseries[,c(2,5)]) %>% 
  mutate(Serie = "Spatial w/ S-R relation mid-strong estimated ")
total6 <- as.data.frame(base.model6$timeseries[,c(2,5)]) %>% 
  mutate(Serie = "Spatial w/ Ricker S-R relation estimated")
total7 <- as.data.frame(base.model7$timeseries[,c(2,5)]) %>% 
  mutate(Serie = "Spatial w/ New set parameters Linf, k and M")

totalbio <- rbind(total01,
              total1,
              total2,
              total3,
              total4,
              total5,
              total6,
              total7)

ggplot(totalbio, aes(x = Yr, y = Bio_all, color=Serie)) +
  geom_point(size=0.7,
             col="black") +
  geom_smooth(method = "loess",
              span=0.75, 
              se=TRUE)+
  labs(title = "Compare variables between scenarios",
       x = "",
       y = "Total biomass (mt)") +
  theme_few() +
  scale_color_viridis_d(option="H",
                        name="Scenarios")+
  theme(axis.text.x = element_text(angle = 90,
                                   hjust = 1))
```

```{r eval=FALSE}
SSplotComps(base.model01,
            subplots = 1,
            maxrows = 5,
            maxcols = 5)
SSplotComps(base.model1,
            subplots = 1,
            maxrows = 5,
            maxcols = 5)
SSplotComps(base.model2,
            subplots = 1,
            maxrows = 5,
            maxcols = 5)
SSplotComps(base.model3,
            subplots = 1,
            maxrows = 5,
            maxcols = 5)
SSplotComps(base.model4,
            subplots = 1,
            maxrows = 5,
            maxcols = 5)
SSplotComps(base.model5,
            subplots = 1,
            maxrows = 5,
            maxcols = 5)
SSplotComps(base.model6,
            subplots = 1,
            maxrows = 5,
            maxcols = 5)
SSplotComps(base.model7,
            subplots = 1,
            maxrows = 5,
            maxcols = 5)
```
Heatmap

```{r}
df <- base.modeltest$lendbase

# Filtra el dataframe
df1 <- df %>% 
  dplyr::filter(Pearson<5) %>% 
  dplyr::select(c(1, 6, 19, 16)) %>% 
  dplyr::mutate(Fleet = dplyr::case_when(
    Fleet == 1 ~ "FISHERYBS",
    Fleet == 2 ~ "FISHERYEI",
    Fleet == 3 ~ "FISHERYGS",
    Fleet == 4 ~ "FISHERYJOIN",
    Fleet == 5 ~ "FISHERYSSIW",
    Fleet == 6 ~ "SURVEYBS",
    Fleet == 7 ~ "SURVEYEI",
    Fleet == 8 ~ "SURVEYGS",
    Fleet == 9 ~ "SURVEYJOIN",
    Fleet == 10 ~ "SURVEYSSIW",
    Fleet == 11 ~ "PREDATOR",
    TRUE ~ NA_character_  # Para manejar cualquier valor que no coincida
  ))

# Crea el heatmap
ggplot(df1, aes(x = Yr, y = Bin, fill = Pearson)) +
  geom_tile() +
  scale_fill_gradient2(low = "blue", 
                       mid = "white", 
                       high = "red",
                       midpoint = 0) +
  facet_wrap(Fleet~., ncol=5) + 
  labs(title = "",
       x = "",
       y = "Length (cm)",
       fill = "Pearson") +
  theme_few()+
  theme(axis.text.x = element_text(angle = 90,
                                   hjust = 1, 
                                   size = 7))

```
hexagon

```{r eval=FALSE}
# Assuming your data is in a data frame called 'lendbase'
# Filter and select relevant columns
data <- base.model1$lendbase %>%
  dplyr::filter(Pearson<5) %>% 
  dplyr::select(c(1, 6, 19, 16)) %>% 
  dplyr::mutate(Fleet = dplyr::case_when(
    Fleet == 1 ~ "FISHERYBS",
    Fleet == 2 ~ "FISHERYEI",
    Fleet == 3 ~ "FISHERYGS",
    Fleet == 4 ~ "FISHERYJOIN",
    Fleet == 5 ~ "FISHERYSSIW",
    Fleet == 6 ~ "SURVEYBS",
    Fleet == 7 ~ "SURVEYEI",
    Fleet == 8 ~ "SURVEYGS",
    Fleet == 9 ~ "SURVEYJOIN",
    Fleet == 10 ~ "SURVEYSSIW",
    Fleet == 11 ~ "PREDATOR",
    TRUE ~ NA_character_  # Para manejar cualquier valor que no coincida
  )) %>% 
  dplyr::select(Yr, Bin, Pearson)

# Create the hexagonal heatmap
ggplot(data, aes(x = Yr, y = Bin)) +
  geom_hex(bins=50) +  # Hexagonal bins
  scale_fill_gradient2(low = "blue", 
                       mid = "white", 
                       high = "red", 
                       midpoint = 0) +  # Color scale
  theme_minimal() +  # Minimal theme for cleaner look
  labs(title = "Hexagonal Heatmap of Pearson Residuals",
       x = "Year",
       y = "Bin",
       fill = "Pearson Residual")  # Labels for plot and axes

ggplot(data, aes(x = Yr, y = Bin)) +
  geom_hex(bins = 50) +
  scale_fill_gradient2(low = "blue", 
                       mid = "white", 
                       high = "red", midpoint = median(data$Bin)) +
  theme_minimal() +
  labs(title = "Hexagonal Heatmap of Pearson Residuals",
       x = "Year", y = "Bin", fill = "Pearson Residual")
```

## Diagnosis Base Model

Step to do a good practice in model diagnosis is;

-     1.  Convergence. Final convergence criteria is 1.0e-04

-     2.  Residual (visual and metrics)

-     3.  Retrospective analysis (Mhon Parameter)

-    4.  Likelihood profile

all this framework try to follow recommendations of @Carvalho2021b

## Residual consistency

```{r}
SSplotRunstest(base.model1.1,
               subplots = "len",
               add=T,
               plot = TRUE,
               plotdir = Figs)
SSplotRunstest(base.model1.2,
               subplots = "len",
               add=T,
               plot = TRUE,
               plotdir = Figs)
```

```{r eval=FALSE}
SSplotRunstest(base.model01,
               subplots = "len",
               add=T)
SSplotRunstest(base.model1,
               subplots = "len",
               add=T)
SSplotRunstest(base.model2,
               subplots = "len",
               add=T)
SSplotRunstest(base.model3,
               subplots = "len",
               add=T)
SSplotRunstest(base.model4,
               subplots = "len",
               add=T)
SSplotRunstest(base.model5,
               subplots = "len",
               add=T)
SSplotRunstest(base.model6,
               subplots = "len",
               add=T)
SSplotRunstest(base.modeltest,
               subplots = "len",
               add=T)
```

```{r}
SSplotRunstest(base.model1.1,
               subplots = "cpue",
               add=T)
SSplotRunstest(base.model1.2,
               subplots = "cpue",
               add=T)
```

```{r eval=FALSE}
SSplotRunstest(base.model01,
               subplots = "cpue",
               add=T)
SSplotRunstest(base.model1,
               subplots = "cpue",
               add=T)
SSplotRunstest(base.model2,
               subplots = "cpue",
               add=T)
SSplotRunstest(base.model3,
               subplots = "cpue",
               add=T)
SSplotRunstest(base.model4,
               subplots = "cpue",
               add=T)
SSplotRunstest(base.model5,
               subplots = "cpue",
               add=T)
SSplotRunstest(base.model6,
               subplots = "cpue",
               add=T)
SSplotRunstest(base.modeltest,
               subplots = "cpue",
               add=T)
```

```{r eval=FALSE}
SSplotJABBAres(base.model01,
               subplots = "len",
               add=T)
SSplotJABBAres(base.model1,
               subplots = "len",
               add=T)
SSplotJABBAres(base.model2,
               subplots = "len",
               add=T)
SSplotJABBAres(base.model3,
               subplots = "len",
               add=T)
SSplotJABBAres(base.model4,
               subplots = "len",
               add=T)
SSplotJABBAres(base.model5,
               subplots = "len",
               add=T)
SSplotJABBAres(base.model6,
               subplots = "len",
               add=T)
SSplotJABBAres(base.model1.2,
               subplots = "len",
               add=T)
```

```{r}
SSplotJABBAres(base.model01,
               subplots = "cpue",
               add=T)
SSplotJABBAres(base.model1,
               subplots = "cpue",
               add=T)
SSplotJABBAres(base.model2,
               subplots = "cpue",
               add=T)
SSplotJABBAres(base.model3,
               subplots = "cpue",
               add=T)
SSplotJABBAres(base.model4,
               subplots = "cpue",
               add=T)
SSplotJABBAres(base.model5,
               subplots = "cpue",
               add=T)
SSplotJABBAres(base.model6,
               subplots = "cpue",
               add=T)
SSplotJABBAres(base.model1.2,
               subplots = "cpue",
               add=T)
```
## Relationship Stock-Recruit

```{r eval=FALSE}
l <- base.modeltest$recruit$SpawnBio
m <- base.modeltest$recruit$pred_recr
ggplot() +
  aes(l, m) +
  geom_point() +
  geom_smooth(method = "loess",
              se = TRUE, 
              color = "blue") +
  theme_few() +
  ylim(0, max(base.modeltest$recruit$pred_recr))

```

## Retrospective analysis

Los análisis retrospectivo, dan cuenta de diferencias de estimación
(sub - sobreestimación) en los patrones entre modelos evaluados.

```{r eval=FALSE}
#one by one
retro(
    dir = dir1.1,
    oldsubdir = "",
    newsubdir = "Retrospective",
    years = 0:-4,
    exe = "ss_osx",
    extras = "-nox",
    skipfinished = FALSE)
```

```{r eval=FALSE}
for (dir in directorios) {
  retro(
    dir = dir,
    oldsubdir = "",
    newsubdir = "Retrospective",
    years = 0:-5,
    exe = "ss_osx",
    extras = "-nox",
    skipfinished = FALSE
  )
}
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
#s01
retroModels01 <- SSgetoutput(dirvec=file.path(dir01,
                                            "Retrospective",
                                            paste("retro",0:-5,
                                                  sep="")))

retroSummary01 <- SSsummarize(retroModels01)
# s1
retroModels1 <- SSgetoutput(dirvec=file.path(dir1,
                                            "Retrospective",
                                            paste("retro",0:-5,
                                                  sep="")))

retroSummary1 <- SSsummarize(retroModels1)
#s2
retroModels2 <- SSgetoutput(dirvec=file.path(dir2,
                                            "Retrospective",
                                            paste("retro",0:-5,
                                                  sep="")))

retroSummary2 <- SSsummarize(retroModels2)
#s3
retroModels3 <- SSgetoutput(dirvec=file.path(dir3,
                                            "Retrospective",
                                            paste("retro",0:-5,
                                                  sep="")))

retroSummary3 <- SSsummarize(retroModels3)
#s4
retroModels4 <- SSgetoutput(dirvec=file.path(dir4,
                                            "Retrospective",
                                            paste("retro",0:-5,
                                                  sep="")))

retroSummary4 <- SSsummarize(retroModels4)
#s5
retroModels5 <- SSgetoutput(dirvec=file.path(dir5,
                                            "Retrospective",
                                            paste("retro",0:-5,
                                                  sep="")))

retroSummary5 <- SSsummarize(retroModels5)
#s6
retroModels6 <- SSgetoutput(dirvec=file.path(dir6,
                                            "Retrospective",
                                            paste("retro",0:-5,
                                                  sep="")))

retroSummary6 <- SSsummarize(retroModels6)
#s7
retroModels7 <- SSgetoutput(dirvec=file.path(dir7,
                                            "Retrospective",
                                            paste("retro",0:-5,
                                                  sep="")))

retroSummary7 <- SSsummarize(retroModels7)
```

```{r eval=FALSE, message=FALSE, warning=FALSE}
#stest
retroModels1.1 <- SSgetoutput(dirvec=file.path(dir1.1,
                                            "Retrospective",
                                            paste("retro",0:-4,
                                                  sep="")))

retroSummary1.1 <- SSsummarize(retroModels1.1)

```

```{r echo=FALSE}
#save(retroSummary, retroModels, file="retro5.Rdata")

retro01 <- SSplotRetro(retroSummary01,
            add=T,
            forecast = F,
            legend = T,
            verbose=F)
retro1 <- SSplotRetro(retroSummary1,
            add=T,
            forecast = F,
            legend = T,
            verbose=F)
retro2 <- SSplotRetro(retroSummary2,
            add=T,
            forecast = F,
            legend = T,
            verbose=F)
retro3 <- SSplotRetro(retroSummary3,
            add=T,
            forecast = F,
            legend = T,
            verbose=F)
retro4 <- SSplotRetro(retroSummary4,
            add=T,
            forecast = F,
            legend = T,
            verbose=F)
retro5 <- SSplotRetro(retroSummary5,
            add=T,
            forecast = F,
            legend = T,
            verbose=F)

retro6 <- SSplotRetro(retroSummary6,
            add=T,
            forecast = F,
            legend = T,
            verbose=F)

retro7 <- SSplotRetro(retroSummary7,
            add=T,
            forecast = F,
            legend = T,
            verbose=F)

retro1.1 <- SSplotRetro(retroSummary1.1,
            add=T,
            forecast = F,
            legend = T,
            verbose=F,
            subplots = c("SSB"))
```

## Hindcast Cross-Validation and prediction skill

Implementing the Hindcast Cross-Validation (HCxval) diagnostic in Stock
Synthesis requires the same model outputs generated by
`r4ss:SS_doRetro()`. As a robust measure of prediction skill, we
implemented the mean absolute scaled error (MASE). In brief, the MASE
score scales the mean absolute. Regarding (A MASE score \> 1 indicates
that the average model forecasts are worse than a random walk.
Conversely, a MASE score of 0.5 indicates that the model forecasts twice
as accurately as a naïve baseline prediction; thus, the model has
prediction skill.

```{r}
hci = SSplotHCxval(retroSummarytest, 
                   add = T, 
                   verbose = F, 
                   legendcex = 0.7)
```

```{r eval=FALSE}
hci = SSplotHCxval(retroSummary01, 
                   add = T, 
                   verbose = F, 
                   legendcex = 0.7)
hci = SSplotHCxval(retroSummary1, 
                   add = T, 
                   verbose = F, 
                   legendcex = 0.7)
hci = SSplotHCxval(retroSummary2, 
                   add = T, 
                   verbose = F, 
                   legendcex = 0.7)
hci = SSplotHCxval(retroSummary3, 
                   add = T, 
                   verbose = F, 
                   legendcex = 0.7) 
hci = SSplotHCxval(retroSummary4, 
                   add = T, 
                   verbose = F, 
                   legendcex = 0.7)
hci = SSplotHCxval(retroSummary5, 
                   add = T, 
                   verbose = F, 
                   legendcex = 0.7)
hci = SSplotHCxval(retroSummary6, 
                   add = T, 
                   verbose = F, 
                   legendcex = 0.7)
hci = SSplotHCxval(retroSummary7, 
                   add = T, 
                   verbose = F, 
                   legendcex = 0.7)
```

## Kobe (status)

```{r}
mvln = SSdeltaMVLN(base.modeltest,
                   Fref = "Btgt", 
                   plot = TRUE,
                   addprj=TRUE,
                   virgin = TRUE,
                   mc=100)

```

```{r eval=FALSE}
mvln = SSdeltaMVLN(base.model01,
                   Fref = "Btgt", 
                   plot = TRUE,
                   addprj=TRUE,
                   virgin = TRUE,
                   mc=100)

mvln = SSdeltaMVLN(base.model1,
                   Fref = "Btgt", 
                   plot = TRUE,
                   addprj=TRUE,
                   virgin = TRUE,
                   mc=100)

mvln = SSdeltaMVLN(base.model2,
                   Fref = "Btgt", 
                   plot = TRUE,
                   addprj=TRUE,
                   virgin = TRUE,
                   mc=100)

mvln = SSdeltaMVLN(base.model3,
                   Fref = "Btgt", 
                   plot = TRUE,
                   addprj=TRUE,
                   virgin = TRUE,
                   mc=100)

mvln = SSdeltaMVLN(base.model4,
                   Fref = "Btgt", 
                   plot = TRUE,
                   addprj=TRUE,
                   virgin = TRUE,
                   mc=100)

mvln = SSdeltaMVLN(base.model5,
                   Fref = "Btgt", 
                   plot = TRUE,
                   addprj=TRUE,
                   virgin = TRUE,
                   mc=100)

mvln = SSdeltaMVLN(base.model6,
                   Fref = "Btgt", 
                   plot = TRUE,
                   addprj=TRUE,
                   virgin = TRUE,
                   mc=100)

mvln = SSdeltaMVLN(base.model1.1,
                   Fref = "Btgt", 
                   plot = TRUE,
                   addprj=TRUE,
                   virgin = TRUE,
                   mc=100)

```

another

```{r eval=FALSE}

nanios<-base.model1$Kobe$Yr
SSBDep<-base.model1$Kobe$B.Bmsy
y_BDPR<-base.model1$Kobe$F.Fmsy

# Integrated Model

kobe1 <- ggplot(base.model1$Kobe)+
  #geom_line(aes(kobebro$BBMSY, kobebro$FFMSY)) +
  geom_rect(aes(xmin = 0, xmax = 0.5, ymin = 0, ymax = 5),
            fill = "#E43338", alpha = 0.5) +
  geom_rect(aes(xmin = 0.5, xmax = 0.75, ymin = 0, ymax = 5),
            fill = "#F2ED23", alpha = 0.5) +
  geom_rect(aes(xmin = 0.75, xmax = 1.25, ymin = 0, ymax = 5),
            fill = "#ACC39A", alpha = 0.5) +
  geom_rect(aes(xmin = 1.25, xmax = 9, ymin = 0, ymax = 1),
            fill = "#608D68", alpha = 0.5) +
  geom_rect(aes(xmin = 1.25, xmax = 9, ymin = 1, ymax = 5),
            fill = "#808080", alpha = 0.5) +
  geom_path(aes(x=base.model1$Kobe$B.Bmsy,y=base.model1$Kobe$F.Fmsy,label=base.model1$Kobe$Yr))+
  geom_point(aes(base.model1$Kobe$B.Bmsy, base.model1$Kobe$F.Fmsy),
             lwd=2) +
  geom_hline(yintercept = 1) +
  geom_vline(xintercept = c(0.5, 0.75, 1.75, 1, 1.25), linetype=2)+
  theme_few()+
  labs(x = expression("BD/BD"[RMS]), y = expression("F/F"[RMS]))+
  # geom_text(data = texto_coords, aes(x = x, y = y, label = etiqueta),
  #            vjust = -0.5)+
  geom_text(aes(x=base.model1$Kobe$B.Bmsy,y=base.model1$Kobe$F.Fmsy,label=base.model1$Kobe$Yr),
            nudge_y = 0.1,size = 3,
            check_overlap = TRUE)
kobe1

```

```{r echo=FALSE}
# model01
tablebias01 <- SShcbias(retroSummary01,quant="SSB",verbose=F)
tablebias01a <- SShcbias(retroSummary01,quant="F",verbose=F)

kbl(tablebias01, booktabs = T,format = "latex",
    caption = "Rho parameter in SSB model s01")  %>% 
    kable_styling(latex_options = "HOLD_position")

kbl(tablebias01a, booktabs = T,format = "latex",
    caption = "Rho parameter in F model s01")  %>% 
    kable_styling(latex_options = "HOLD_position")

# model1
tablebias1 <- SShcbias(retroSummary1,quant="SSB",verbose=F)
tablebias1a <- SShcbias(retroSummary1,quant="F",verbose=F)

kbl(tablebias1, booktabs = T,format = "latex",
    caption = "Rho parameter in SSB model s1")  %>% 
    kable_styling(latex_options = "HOLD_position")

kbl(tablebias1a, booktabs = T,format = "latex",
    caption = "Rho parameter in F model s1")  %>% 
    kable_styling(latex_options = "HOLD_position")

# model2
tablebias2 <- SShcbias(retroSummary2,quant="SSB",verbose=F)
tablebias2a <- SShcbias(retroSummary2,quant="F",verbose=F)

kbl(tablebias2, booktabs = T,format = "latex",
    caption = "Rho parameter in SSB model s2")  %>% 
    kable_styling(latex_options = "HOLD_position")

kbl(tablebias2a, booktabs = T,format = "latex",
    caption = "Rho parameter in F model s2")  %>% 
    kable_styling(latex_options = "HOLD_position")

# model3
tablebias3 <- SShcbias(retroSummary3,quant="SSB",verbose=F)
tablebias3a <- SShcbias(retroSummary3,quant="F",verbose=F)

kbl(tablebias3, booktabs = T,format = "latex",
    caption = "Rho parameter in SSB model s3")  %>% 
    kable_styling(latex_options = "HOLD_position")

kbl(tablebias3a, booktabs = T,format = "latex",
    caption = "Rho parameter in F model s3")  %>% 
    kable_styling(latex_options = "HOLD_position")

# model4
tablebias4 <- SShcbias(retroSummary4,quant="SSB",verbose=F)
tablebias4a <- SShcbias(retroSummary4,quant="F",verbose=F)

kbl(tablebias4, booktabs = T,format = "latex",
    caption = "Rho parameter in SSB model s4")  %>% 
    kable_styling(latex_options = "HOLD_position")

kbl(tablebias4a, booktabs = T,format = "latex",
    caption = "Rho parameter in F model s4")  %>% 
    kable_styling(latex_options = "HOLD_position")

# model5
tablebias5 <- SShcbias(retroSummary5,quant="SSB",verbose=F)
tablebias5a <- SShcbias(retroSummary5,quant="F",verbose=F)

kbl(tablebias5, booktabs = T,format = "latex",
    caption = "Rho parameter in SSB model s5")  %>% 
    kable_styling(latex_options = "HOLD_position")

kbl(tablebias5a, booktabs = T,format = "latex",
    caption = "Rho parameter in F model s5")  %>% 
    kable_styling(latex_options = "HOLD_position")

# model6
tablebias6 <- SShcbias(retroSummary6,quant="SSB",verbose=F)
tablebias6a <- SShcbias(retroSummary6,quant="F",verbose=F)

kbl(tablebias6, booktabs = T,format = "latex",
    caption = "Rho parameter in SSB model s6")  %>% 
    kable_styling(latex_options = "HOLD_position")

kbl(tablebias6a, booktabs = T,format = "latex",
    caption = "Rho parameter in F model s6")  %>% 
    kable_styling(latex_options = "HOLD_position")

# model7
tablebias7 <- SShcbias(retroSummary7,quant="SSB",verbose=F)
tablebias7a <- SShcbias(retroSummary7,quant="F",verbose=F)

kbl(tablebias7, booktabs = T,format = "latex",
    caption = "Rho parameter in SSB model s7")  %>% 
    kable_styling(latex_options = "HOLD_position")

kbl(tablebias7a, booktabs = T,format = "latex",
    caption = "Rho parameter in F model s7")  %>% 
    kable_styling(latex_options = "HOLD_position")

```

## Likelihood tables

```{r}
like01 <- base.model01$likelihoods_used
like01$model <- rep("s01", nrow(like01))
like01 <- rownames_to_column(like01, var = "Description")
like1 <- base.model1$likelihoods_used
like1$model <- rep("s1", nrow(like1))
like1 <- rownames_to_column(like1, var = "Description")
like2 <- base.model1$likelihoods_used
like2$model <- rep("s2", nrow(like2))
like2 <- rownames_to_column(like2, var = "Description")
like3 <- base.model3$likelihoods_used
like3$model <- rep("s3", nrow(like3))
like3 <- rownames_to_column(like3, var = "Description")
like4 <- base.model4$likelihoods_used
like4$model <- rep("s4", nrow(like4))
like4 <- rownames_to_column(like4, var = "Description")
like5 <- base.model5$likelihoods_used
like5$model <- rep("s5", nrow(like5))
like5 <- rownames_to_column(like5, var = "Description")
like6 <- base.model6$likelihoods_used
like6$model <- rep("s6", nrow(like6))
like6 <- rownames_to_column(like6, var = "Description")
like7 <- base.model7$likelihoods_used
like7$model <- rep("s7", nrow(like7))
like7 <- rownames_to_column(like7, var = "Description")

totalike <- rbind(like01,
                  like1,
                   like2,
                   like3,
                   like4,
                   like5,
                   like6,
                   like7)

```

```{r}
ggplot(totalike)+
  geom_bar( aes(y = reorder(Description, values), x =log(values)),
            stat = "identity", position = "dodge") +
  theme_few() +
  facet_wrap(.~model, 
             ncol=4)+
  labs(y="",
       x="Log Likelihood")+
  theme(axis.text.x = element_text(angle = 90, 
                                   hjust = 1))+
  xlim(0,20)
```

## Likelihood Profile

```{r eval=FALSE, echo = FALSE}
# 1. Identificar el directorio donde se encuentra el modelo base ----
dirname.model.run <- here("s1.2")
# 2. Crear un nuevo directorio para el "Perfil_Verosimilitud"  
dirname.R0.profile <- here("s1.2",
                           "Perfil_Verosimilitud")
dir.create(path=dirname.R0.profile, 
           showWarnings = TRUE, 
           recursive = TRUE)
# 3. Crear un subdirectorio llamado "plots_Verosimilitud" ----
plotdir=paste0(dirname.R0.profile, "/plots_Verosimilitud")
dir.create(path=plotdir,
           showWarnings = TRUE, 
           recursive = TRUE)
# 4. Crear un subdirectorio llamado "simple" ----
reference.dir <- paste0(dirname.R0.profile,'/simple') 
dir.create(path=reference.dir, showWarnings = TRUE, recursive = TRUE)
# 5. Copiar el resultado del modelo base completo en este directorio ----
file.copy(Sys.glob(paste(dirname.model.run, "*.*", sep="/"),
                   dirmark = FALSE),
                    reference.dir)
# 6. Leer la salida del modelo base ----
Base <- SS_output(dir=reference.dir,covar=T)
# 7. Copiar los archivos necesarios de "simple" al directorio "Perfil_Verosimilitud" ----
copy_SS_inputs(dir.old = reference.dir, 
               dir.new =  dirname.R0.profile,
               copy_exe = TRUE,
               verbose = FALSE)
# 8. Leer los archivos del modelo ----
inputs <- r4ss::SS_read(dir = dirname.R0.profile)
# 9. Editar el archivo control la fase de estimación recdev ----
inputs$ctl$recdev_phase <- 1
# 10. Editar el archivo starter para leer los valores de inicio ----
inputs$start$init_values_src <- 0
# 11. Vector de valores para el perfil ----
R0.vec <- seq(18,30,1)  
Nprof.R0 <- length(R0.vec)
# 12. Cambiar el nombre del archivo control en el archivo starter.ss ----
inputs$start$ctlfile <- "control_modified.ss" 
# 13. Incluir prior_like para parámetros no estimados ----
inputs$start$prior_like <- 1                                 
# 14. Escribir los modelos modificados ----
r4ss::SS_write(inputs, dir = dirname.R0.profile, overwrite = TRUE)
# 15. Ejecutar la función profile() ----
#?SS_profile()
profile <- profile(dir=dirname.R0.profile, # directory
                      exe="ss_osx",
                      oldctlfile ="control.ss",
                      newctlfile="control_modified.ss",
                      string="SR_LN(R0)",
                      profilevec=R0.vec)
# 16. Leer los archivos de salida ----
# (con nombres como Report1.sso, Report2.sso, etc.)
prof.R0.models <- SSgetoutput(dirvec=dirname.R0.profile, 
                              keyvec=1:Nprof.R0, 
                              getcovar = FALSE) 
# 17. Resumir las salidas con la función SSsummarize()  ----
prof.R0.summary <- SSsummarize(prof.R0.models)
# 18. Identificar los componentes de Verosimilitud ----
mainlike_components         <- c('TOTAL',
                                 "Survey", 
                                 'Length_comp',
                                 "Age_comp",
                                 "Catch",
                                 'Size_at_age',
                                 'Recruitment') 
mainlike_components_labels  <- c('Total likelihood',
                                 'Index likelihood',
                                 'Length likelihood',
                                 "Age likelihood",
                                 "Catch Likelihood",
                                 'Size_at_age likelihood',
                                 'Recruitment likelihood')

```

Funciones para generar plots de perfil de verosimilitud ----

### SSplotProfile() ----

```{r eval=FALSE, echo = FALSE}
png(file.path(plotdir,"R0_profile_plot.png"),
    width=7,
    height=4.5,
    res=300,
    units='in')
par(mar=c(5,4,1,1))
SSplotProfile(prof.R0.summary,           # summary object
              profile.string = "R0",     # substring of profile parameter
              profile.label=expression(log(italic(R)[0])), 
              ymax=2050,minfraction = 0.001,
              pheight=4.5, 
              print=FALSE, 
              plotdir=plotdir, 
              components = mainlike_components, 
              component.labels = mainlike_components_labels,
              add_cutoff = TRUE,
              cutoff_prob = 0.95)

Baseval <- round(Base$parameters$Value[grep("R0",Base$parameters$Label)],2)
Baselab <- paste(Baseval,sep="")
axis(1,at=Baseval,label=Baselab)
abline(v = Baseval, lty=2)
dev.off()
```

## Convergence criteria

0.0001 final convergence criteria (e.g. 1.0e-04)

```{r eval=FALSE, echo = FALSE}
# Comparación de series de tiempo 
labs <- paste("SR_Ln(R0) = ",R0.vec)
labs[which(round(R0.vec,2)==Baseval)] <- paste("SR_Ln(R0) = ",
                                               Baseval,"(Base model)")

SSplotComparisons(prof.R0.summary,
                  legendlabels=labs,
                  pheight=4.5,png=TRUE,
                  plotdir=plotdir,
                  legendloc='bottomleft')

```

piner Plot

```{r echo = FALSE}
#### R0_profile_plot_Length_like ----
png(file.path(plotdir,"R0_profile_plot_Length_like.png"),
    width=7,
    height=4.5,
    res=300,
    units='in')
par(mar=c(5,4,1,1))
PinerPlot(prof.R0.summary, 
          profile.string = "R0", 
          component = "Length_like",
          main = "Changes in length-composition likelihoods by fleet",
          add_cutoff = TRUE,
          cutoff_prob = 0.95)
Baseval <- round(Base$parameters$Value[grep("SR_LN",
                                      Base$parameters$Label)],2)
Baselab <- paste(Baseval,sep="")
axis(1,at=Baseval,
     label=Baselab)
abline(v = Baseval, lty=2)
dev.off()
```

```{r echo = FALSE}
#### R0_profile_plot_Survey_like ----
png(file.path(plotdir,"R0_profile_plot_Survey_like.png"),
    width=7,
    height=4.5,
    res=300,
    units='in')
par(mar=c(5,4,1,1))
PinerPlot(prof.R0.summary, 
          profile.string = "R0", 
          component = "Surv_like",
          main = "Changes in Index likelihoods by fleet",
          add_cutoff = TRUE,
          cutoff_prob = 0.95, legendloc="topleft")
Baseval <- round(Base$parameters$Value[grep("SR_LN",
                                            Base$parameters$Label)],2)
Baselab <- paste(Baseval,sep="")
axis(1,at=Baseval,label=Baselab)
abline(v = Baseval, lty=2)
dev.off()
```

## Outputs Model Base

```{r}
outps1 <- base.modeltest$timeseries[1:53, 2:8]

# addWorksheet(wb, "variable_s2")
# writeData(wb, "variable_s2", outps1)
# 
# # Guardar el workbook
# saveWorkbook(wb, "DataKrill.xlsx", overwrite = TRUE)
# 
# 
# out_s2 <- read_excel("DataKrill.xlsx", 
#     sheet = "variable_s2")
outps1 %>%
  kbl(booktabs = TRUE,
      format = "html",
    caption = "Main variables outputs from stock asssessment krill in WAP") %>%
  kable_paper("hover", 
              full_width = TRUE)%>%
  kable_styling(latex_options = c("striped",
                                  "condensed"),
                full_width = FALSE,
                font_size=9,
                html_font = "arial")#%>% 
  #pack_rows(index = c("Estimation" = 1,
   #                     "Prediction" = 45))
```

## Comparision outputs betwwen scenarios

```{r echo=FALSE}

#read in all model runs
#note if cover=T you need a hessian; if covar=F you do not need a hessian
biglist <- SSgetoutput(keyvec = NULL,
                       dirvec = c(dir01,
                                  dir1,
                                  dir2,
                                  dir3,
                                  dir4,
                                  dir5,
                                  dir6,
                                  dir7),
                       getcovar = F)

#create summary of model runs from list above
summaryoutputall <- SSsummarize(biglist)
```

```{r echo=FALSE}
compall <- SSplotComparisons(summaryoutputall,
                  legendlabels = c("No Spatial",
                                   "s1",
                                   "s2",
                 "s3",
                 "s4",
                 "s5",
                 "s6",
                 "s7"),
                 labels = c("Year", 
                            "Spawning biomass (t)",
                            "Relative spawning biomass", 
                            "Age-0 recruits (1,000s)",
                            "Recruitment deviations", 
                            "Index", "Log index", 
                            "1 - SPR", 
                            "Density",
                            "Management target", 
                            "Minimum stock size threshold",
                            "Spawning output",
                            "Harvest rate"),
                 png=TRUE,plotdir=Figs)

```

```{r}
comtable <- SStableComparisons(summaryoutputall,
                   likenames = c("TOTAL", 
                                 "Survey", 
                                 "Length_comp",
                                 "Age_comp", 
                                 "priors",
                                 "Size_at_age"), 
                   names = c("Recr_Virgin",
                             "R0", 
                             "steep",
                             "NatM",
                             "L_at_Amax", 
                             "VonBert_K", 
                             "SSB_Virg", 
                             "Bratio_2020",
                             "SPRratio_2020"),
                   digits = NULL,
                   modelnames = c("s01", "s1","s2",
                 "s3",
                 "s4",
                 "s5",
                 "s6",
                 "s7"),
                   csv = F,
                   csvdir =~"/DOCAS/SA_Krill",
                   csvfile = "parameter_comparison_table.csv",
                 verbose = TRUE,
                   mcmc = FALSE)
kbl(comtable, booktabs = T,format = "latex",
    caption = "Comparacion likelihood y parámetros s01, s1, s2, s3, s4, s5, s6 y s7")  %>% 
    kable_styling(latex_options = "scale_down")

```

## Comparision between select models `No spatial`and `Spatial implicit` and `Spatial W/ new set parametres`

```{r echo=FALSE}
#read in all model runs
#note if cover=T you need a hessian; if covar=F you do not need a hessian
biglist1_2_3 <- SSgetoutput(keyvec = NULL,
                       dirvec = c(
                                  dir1.1,
                                  dir1.2,
                                  dir1.3),
                       getcovar = F)

#create summary of model runs from list above
summaryoutput1_2_3 <- SSsummarize(biglist1_2_3)

SSplotComparisons(summaryoutput1_2_3,
                  legendlabels = c("Ref Model: No Env-Predator",
                 "S1.1 w/ Predator data",
                 "S1.1 w/ Env and Predator data"),
                 filenameprefix = "COM1",
                 labels = c("Year", 
                            "Spawning biomass (t)",
                            "Relative spawning biomass", 
                            "Age-0 recruits (1,000s)",
                            "Recruitment deviations", 
                            "Index", "Log index", 
                            "1 - SPR", 
                            "Density",
                            "Management target", 
                            "Minimum stock size threshold",
                            "Spawning output",
                            "Harvest rate"),
                 png=TRUE,
                 plotdir=Figs)
```

comparision between select models `Old Paramters` and `New Parameters WG SAM 2024/23`

```{r echo=FALSE}
#read in all model runs
#note if cover=T you need a hessian; if covar=F you do not need a hessian
biglist1_7 <- SSgetoutput(keyvec = NULL,
                       dirvec = c(
                                  dir1,
                                  dir7),
                       getcovar = F)

#create summary of model runs from list above
summaryoutput1_7 <- SSsummarize(biglist1_7)

comp1_7 <- SSplotComparisons(summaryoutput1_7,
                  legendlabels = c(
                                   "Old Parameters",

                 "Parameters WG SAM 2024/27"),
                 labels = c("Year", 
                            "Spawning biomass (t)",
                            "Relative spawning biomass", 
                            "Age-0 recruits (1,000s)",
                            "Recruitment deviations", 
                            "Index", "Log index", 
                            "1 - SPR", 
                            "Density",
                            "Management target", 
                            "Minimum stock size threshold",
                            "Spawning output",
                            "Harvest rate"),
                 png=TRUE,plotdir=Figs)
```
### Comparision between select models `s1` and `s9`

```{r echo=FALSE}
#read in all model runs
#note if cover=T you need a hessian; if covar=F you do not need a hessian
biglist1_9 <- SSgetoutput(keyvec = NULL,
                       dirvec = c(
                                  dir1,
                                  dir9),
                       getcovar = F)

#create summary of model runs from list above
summaryoutput1_9 <- SSsummarize(biglist1_9)

comp1_9 <- SSplotComparisons(summaryoutput1_9,
                  legendlabels = c(
                                   "Base",
                 "selectivity Block 2006"),
                 filenameprefix = "COM1_9",
                 labels = c("Year", 
                            "Spawning biomass (t)",
                            "Relative spawning biomass", 
                            "Age-0 recruits (1,000s)",
                            "Recruitment deviations", 
                            "Index", "Log index", 
                            "1 - SPR", 
                            "Density",
                            "Management target", 
                            "Minimum stock size threshold",
                            "Spawning output",
                            "Harvest rate"),
                 png=TRUE,plotdir=Figs)
```
## Comparsion in sd long term time series

```{r fig.height=6, fig.width=8}

sblt01 <- base.model01$timeseries$Bio_all[1:50]
sblt1 <-  base.model1$timeseries$Bio_all
sblt2 <-  base.model2$timeseries$Bio_all
sblt3 <-  base.model3$timeseries$Bio_all
sblt4 <-  base.model4$timeseries$Bio_all
sblt5 <-  base.model5$timeseries$Bio_all
sblt6 <-  base.model6$timeseries$Bio_all
sblt7 <-  base.model7$timeseries$Bio_all

errt <- as.data.frame(cbind(sblt01,
              sblt1,
              sblt2,
              sblt3,
              sblt4,
              sblt5,
              sblt6,
              sblt7))
err_long <- errt %>%
  pivot_longer(cols = everything(),
               names_to = "variable",
               values_to = "value")  
titles <- c("No Spatial",
            "Spatial implicit", 
            "Spatial w/ without S-R relation",
            "Spatial w/ S-R relation weak (0.9 steepness)",
            "Spatial w/ S-R relation strong (0.6 steepness)",
            "Spatial w/ S-R relation mid-strong estimated",
            "Spatial w/ Ricker S-R relation estimated", 
            "Spatial w/ New set parameters Linf, k and M")

ggplot(err_long, aes(x = variable, y = value)) +
  geom_boxplot() +
  geom_jitter(width = 0.2, alpha = 0.3) +
  labs(title = "Long Term Predictions",
       x = "",
       y = "Spawning Biomass (t)") +
  theme_few() +
  scale_x_discrete(labels= titles)+
  theme(axis.text.x = element_text(angle = 90,
                                   hjust = 1))
```

## Autocorrelation in Recruit and Biomas

```{r fig.width=8, fig.height=7}
recs01 <- base.model01$recruit$pred_recr
recs1 <- base.model1$recruit$pred_recr
recs2 <- base.model2$recruit$pred_recr
recs3 <- base.model3$recruit$pred_recr
recs4 <- base.model4$recruit$pred_recr
recs5 <- base.model5$recruit$pred_recr
recs6 <- base.model6$recruit$pred_recr
recs7 <- base.model7$recruit$pred_recr

p01 <- ggAcf(base.model01$recruit$pred_recr) +
  ggtitle("No Spatial") +
  theme_few()
p1 <- ggAcf(base.model1$recruit$pred_recr) +
  ggtitle("Spatial implicit") +
  theme_few()
p2 <- ggAcf(base.model2$recruit$pred_recr) +
  ggtitle("Spatial w/ without S-R relation") +
  theme_few()
p3 <- ggAcf(base.model3$recruit$pred_recr) +
  ggtitle("Spatial w/ S-R relation weak (0.9 steepness)") +
  theme_few()
p4 <- ggAcf(base.model4$recruit$pred_recr) +
  ggtitle("Spatial w/ S-R relation strong (0.6 steepness)") +
  theme_few()
p5 <- ggAcf(base.model5$recruit$pred_recr) +
  ggtitle("Spatial w/ S-R relation mid-strong estimated") +
  theme_few()
p6 <- ggAcf(base.model6$recruit$pred_recr) +
  ggtitle("Spatial w/ Ricker S-R relation estimated") +
  theme_few()
p7 <- ggAcf(base.model7$recruit$pred_recr) +
  ggtitle("Spatial w/ New set parameters Linf, k and M") +
  theme_few()

ggarrange(p01,p1,
          p2,
          p3,
          p4,
          p5,
          p6,
          p7, ncol=3, nrow = 3,
          common.legend = T)

```

## Recruit deviation

```{r}
dev01 <- base.model01$recruit[1:43,c(1,7)] %>% 
  mutate(Serie = "S01")
dev1 <- base.model1$recruit[1:43,c(1,7)]%>% 
  mutate(Serie = "S1")
dev2 <- base.model2$recruit[1:43,c(1,7)] %>% 
  mutate(Serie = "S2")
dev3 <- base.model3$recruit[1:43,c(1,7)] %>% 
  mutate(Serie = "S3")
dev4 <- base.model4$recruit[1:43,c(1,7)] %>% 
  mutate(Serie = "S4")
dev5 <- base.model5$recruit[1:43,c(1,7)] %>% 
  mutate(Serie = "S5")
dev6 <- base.model6$recruit[1:43,c(1,7)] %>% 
  mutate(Serie = "S6")
dev7 <- base.model7$recruit[1:43,c(1,7)] %>% 
  mutate(Serie = "S7")

data_list <- list(dev01, dev1, dev2, dev3, dev4, dev5, dev6, dev7)
titles <- c("No Spatial",
            "Spatial implicit", 
            "Spatial w/ without S-R relation",
            "Spatial w/ S-R relation weak (0.9 steepness)",
            "Spatial w/ S-R relation strong (0.6 steepness)",
            "Spatial w/ S-R relation mid-strong estimated",
            "Spatial w/ Ricker S-R relation estimated", 
            "Spatial w/ New set parameters Linf, k and M")
# Lista para almacenar los gráficos
plots <- list()

for (i in 1:length(data_list)) {
  p <- ggplot(data_list[[i]], aes(x = Yr, y = dev)) +
    geom_point() +
    stat_smooth(method = "loess", 
                span = 0.15, 
                se = T) +
    geom_hline(yintercept = 0,
               col="red",
               linetype = "dotdash")+
    ggtitle(titles[i]) +
    theme(axis.text.x = element_text(angle = 90,
                                   hjust = 1))+
    theme_few()
  plots[[i]] <- p
}

# Organizar los gráficos con ggarrange
combined_plot <- ggarrange(plotlist = plots, ncol = 3, nrow = 3)

combined_plot
```
## Platoons analisis

```{r}
# Definir los valores del eje x
x_values <- seq(0, 7, by=0.1)

# Definir las medias y desviaciones estándar para los valores específicos
means <- c( 2.9,  3.5, 4.1)
std_devs <- c(0.65, 0.8, 0.65)

# Crear un dataframe que contenga los valores de x, la media y las desviaciones estándar
data <- data.frame(
  x = rep(x_values, each = length(means)),
  mean = rep(means, times = length(x_values)),
  std_dev = rep(std_devs, times = length(x_values))
)

# Calcular las curvas de las desviaciones estándar
data <- data %>%
  mutate(y = exp(-((x - mean)^2) / (2 * std_dev^2)))

# Graficar con ggplot
ggplot(data, aes(x = x, y = y, linetype = as.factor(mean))) +
  geom_line(size = 1) +
  scale_linetype_manual(values = means,
                        name ="Platoon") +
  labs(x = "", y = "growth increment") +
  theme_minimal()+
  xlim(0,7)

```
## AKL

In a catch-at-length model like krill assessment the AKL matrix is modelled trought parametrization process

```{r AKL, out.width='60%',  fig.cap="Representation of ALK Matrix to krill in 48.1"}
alk_matrix <- base.model1$ALK[,,1] # Ajusta las dimensiones según tu matriz

# Convertir la matriz en un data.frame
alk_df <- as.data.frame(as.table(alk_matrix))

# Renombrar columnas para mayor claridad
colnames(alk_df) <- c("Length", "TrueAge", "Value")

# Asegúrate de que las columnas Length y TrueAge sean numéricas
alk_df$Length <- as.numeric(as.character(alk_df$Length))
alk_df$TrueAge <- as.numeric(as.character(alk_df$TrueAge))

# Asegúrate de que la columna Value sea numérica
alk_df$Value <- as.numeric(alk_df$Value)

# Crear el plot usando ggplot
ggplot(alk_df, aes(x = TrueAge, y = Length, fill = Value)) +
  geom_tile() +
  scale_fill_gradient(low = "white", high = "blue") +
  labs(x = "True Age",
       y = "Length",
       fill = "Value") +
  theme_bw()+
  theme(legend.position = "none")
```

\newpage

# REFERENCES