### 2025-09-16 Supp_Mat_2.Rmd

---
title: "Supplementary Material 2"
subtitle: "Mathematical Formulas in SS3 for Krill Stock Assessment"
output: 
  pdf_document:
    latex_engine: xelatex
    toc: true
    toc_depth: 3
    number_sections: false
geometry: margin=2cm
fontsize: 11pt
header-includes:
  - \usepackage{amsmath}
  - \usepackage{amssymb}
  - \usepackage{bm}
classoption: fleqn
---

\pagebreak

### State Model

- Population Dynamics

\begin{equation}
N_{a,l,s,t+1} = N_{a-1,l,s,t} \cdot e^{-(M_{a-1,l,s,t} + F_{a-1,l,s,t})} \cdot T_{a-1 \rightarrow a,l,s,t} + R_{a,l,s,t}
\end{equation}

- Total Mortality

\begin{equation}
Z_{a,l,s,t} = M_{a,l,s,t} + F_{a,l,s,t}
\end{equation}

- Spawning Stock Biomass

\begin{equation}
SSB_t = \sum_a \sum_l \sum_s N_{a,l,s,t} \cdot W_{a,l,s,t} \cdot \text{Mat}_{a,l,s,t} \cdot \text{Fec}_{a,l,s,t}
\end{equation}

- Total Biomass

\begin{equation}
B_t = \sum_a \sum_l \sum_s N_{a,l,s,t} \cdot W_{a,l,s,t}
\end{equation}

### Sub-models

- Selectivity

\begin{equation}
S_{a,l,f} = \frac{1}{1 + e^{-\ln(19) \cdot \frac{(L_l - L_{50,f})}{(L_{95,f} - L_{50,f})}}}
\end{equation}

\begin{equation}
S_{a,f,t} = \sum_l S_{l,f,t} \cdot P(L_l|a,t)
\end{equation}

- Stock-Recruitment

\begin{equation}
R_t = \frac{\alpha \cdot SSB_{t-\tau}}{\beta + SSB_{t-\tau}} \cdot e^{\delta_t - \frac{\sigma_R^2}{2}}
\end{equation}

- Environmental Effects on Recruitment

\begin{equation}
E(\text{Recruitment}) = f(\text{SSB}) \times \exp(B \times \text{env}) \times \exp\left(-0.5 \times \pi_R^2\right)
\end{equation}

- Maturity

\begin{equation}
\text{Mat}_{l} = \frac{1}{1 + e^{-\ln(19) \cdot \frac{(L_l - L_{50})}{(L_{95} - L_{50})}}}
\end{equation}

\begin{equation}
\text{Mat}_{a,t} = \sum_l \text{Mat}_l \cdot P(L_l|a,t)
\end{equation}

- Growth (Von Bertalanffy)

\begin{equation}
L_{a,s} = L_{\infty,s} \cdot (1 - e^{-K_s \cdot (a - t_{0,s})})
\end{equation}

- Weight-Length Relationship

\begin{equation}
W_{l,s} = w_s \cdot L_l^{b_s}
\end{equation}

- Length-at-Age Variability

\begin{equation}
P(L_l|a,s,t) = \frac{1}{\sqrt{2\pi}\sigma_{L|a,s}} \exp\left(-\frac{(L_l - \bar{L}_{a,s,t})^2}{2\sigma_{L|a,s}^2}\right)
\end{equation}

- Length-to-Age Transition Matrix

\begin{equation}
T_{a,l \rightarrow a+1,l',s,t} = P(L_{l'}|a+1,s,t+1) \cdot \text{Growth}_{l \rightarrow l'}
\end{equation}

- Fishing Mortality

\begin{equation}
F_{a,l,s,f,t} = F'_{f,t} \cdot S_{a,l,s,f,t}
\end{equation}

\begin{equation}
F_{a,l,s,t} = \sum_f F_{a,l,s,f,t}
\end{equation}

- Predator Mortality

\begin{equation}
M_{a,t} = M1_{a,t} + M2_{a,t}
\end{equation}

\begin{equation}
M2_{a,t} = \sum_{p=1}^{n_{\text{pred}}} \frac{C_{p,t} \cdot S_{p,a}}{B_{\text{avail},t}}
\end{equation}

### Catch Equations

\begin{equation}
C_{a,l,s,f,t} = \frac{F_{a,l,s,f,t}}{Z_{a,l,s,t}} \cdot N_{a,l,s,t} \cdot (1 - e^{-Z_{a,l,s,t}}) \cdot W_{a,l,s,t}
\end{equation}

\begin{equation}
C_{f,t} = \sum_a \sum_l \sum_s C_{a,l,s,f,t}
\end{equation}

### Observation Model

\begin{equation}
\hat{I}_{q,t} = Q_q \cdot \sum_a \sum_l \sum_s N_{a,l,s,t} \cdot S_{a,l,s,q,t} \cdot W_{a,l,s,t}
\end{equation}

\begin{equation}
\hat{P}_{a,f,t} = \frac{C_{a,f,t}}{\sum_a C_{a,f,t}}
\end{equation}

\begin{equation}
\hat{P}_{l,f,t} = \frac{\sum_a C_{a,l,f,t}}{\sum_a \sum_l C_{a,l,f,t}}
\end{equation}

### Likelihood Components

\begin{equation}
L_{\text{survey}} = -\sum_q \sum_t \frac{1}{2} \left[ \frac{(\ln(I_{q,t}) - \ln(\hat{I}_{q,t}))^2}{\sigma_{I,q}^2} + \ln(2\pi\sigma_{I,q}^2) \right]
\end{equation}

\begin{equation}
L_{\text{age}} = -\sum_f \sum_t N_{eff,f,t} \sum_a P_{a,f,t} \ln(\hat{P}_{a,f,t})
\end{equation}

\begin{equation}
L_{\text{length}} = -\sum_f \sum_t N_{eff,f,t} \sum_l P_{l,f,t} \ln(\hat{P}_{l,f,t})
\end{equation}

\begin{equation}
L_{\text{rec}} = -\frac{1}{2} \sum_t \left[ \frac{\delta_t^2}{\sigma_R^2} + \ln(2\pi\sigma_R^2) \right]
\end{equation}

### Parameter Estimation

\begin{equation}
-\ln(L_{\text{total}}) = -\sum_i w_i \ln(L_i)
\end{equation}

\begin{equation}
\text{Penalty} = \lambda \sum_j \left( \frac{\theta_j - \theta_{j,\text{prior}}}{\sigma_{\theta,j}} \right)^2
\end{equation}

\pagebreak

### Glossary of Terms and Parameters

| **Symbol** | **Description** | **Units** |
|------------|-----------------|-----------|
| $N_{a,l,s,t}$ | Number of individuals at age $a$, length $l$, sex $s$, time $t$ | Thousands |
| $M_{a,l,s,t}$ | Natural mortality rate | yr$^{-1}$ |
| $F_{a,l,s,t}$ | Fishing mortality rate | yr$^{-1}$ |
| $Z_{a,l,s,t}$ | Total mortality rate | yr$^{-1}$ |
| $R_t$ | Recruitment at time $t$ | Thousands |
| $SSB_t$ | Spawning stock biomass at time $t$ | Metric tons |
| $B_t$ | Total biomass at time $t$ | Metric tons |
| $W_{a,l,s,t}$ | Weight-at-age-length-sex | Grams |
| $\text{Mat}_{a,l,s,t}$ | Proportion mature at age-length-sex | Proportion |
| $\text{Fec}_{a,l,s,t}$ | Fecundity-at-age-length-sex | Eggs/individual |
| $S_{a,l,f}$ | Selectivity for fleet $f$ at age $a$, length $l$ | Proportion |
| $L_{\infty,s}$ | Asymptotic length for sex $s$ | mm |
| $K_s$ | Growth rate parameter for sex $s$ | yr$^{-1}$ |
| $t_{0,s}$ | Theoretical age at length zero for sex $s$ | Years |
| $w_s$ | Weight-length scaling parameter for sex $s$ | g/mm$^b$ |
| $b_s$ | Weight-length allometric exponent for sex $s$ | Dimensionless |
| $L_{50}$ | Length at 50% maturity | mm |
| $L_{95}$ | Length at 95% maturity | mm |
| $L_{50,f}$ | Length at 50% selectivity for fleet $f$ | mm |
| $L_{95,f}$ | Length at 95% selectivity for fleet $f$ | mm |
| $\alpha$ | Stock-recruitment steepness parameter | Various |
| $\beta$ | Stock-recruitment capacity parameter | Various |
| $\sigma_R$ | Recruitment variability | Dimensionless |
| $\delta_t$ | Recruitment deviation in year $t$ | Log-scale |
| $Q_q$ | Catchability coefficient for survey $q$ | Various |
| $C_{a,l,s,f,t}$ | Catch at age-length-sex by fleet $f$ at time $t$ | Metric tons |
| $I_{q,t}$ | Observed survey index for survey $q$ at time $t$ | Various |
| $\hat{I}_{q,t}$ | Expected survey index for survey $q$ at time $t$ | Various |
| $P_{a,f,t}$ | Observed age composition for fleet $f$ at time $t$ | Proportion |
| $\hat{P}_{a,f,t}$ | Expected age composition for fleet $f$ at time $t$ | Proportion |
| $P_{l,f,t}$ | Observed length composition for fleet $f$ at time $t$ | Proportion |
| $\hat{P}_{l,f,t}$ | Expected length composition for fleet $f$ at time $t$ | Proportion |
| $N_{eff,f,t}$ | Effective sample size for fleet $f$ at time $t$ | Dimensionless |
| $\sigma_{I,q}$ | Standard deviation of survey index $q$ | Log-scale |
| $M1_{a,t}$ | Base natural mortality at age $a$, time $t$ | yr$^{-1}$ |
| $M2_{a,t}$ | Predation mortality at age $a$, time $t$ | yr$^{-1}$ |
| $C_{p,t}$ | Consumption by predator $p$ at time $t$ | Metric tons |
| $S_{p,a}$ | Age-specific selectivity by predator $p$ | Proportion |
| $B_{\text{avail},t}$ | Available biomass for predation at time $t$ | Metric tons |
| $\beta_i$ | Environmental coefficient for covariate $i$ | Various |
| $X_{i,t}$ | Environmental covariate $i$ at time $t$ | Various |
| $\lambda$ | Penalty weight | Dimensionless |
| $w_i$ | Likelihood component weight | Dimensionless |
| $\theta_j$ | Estimated parameter $j$ | Various |
| $\theta_{j,\text{prior}}$ | Prior mean for parameter $j$ | Various |
| $\sigma_{\theta,j}$ | Prior standard deviation for parameter $j$ | Various |
| $\tau$ | Recruitment lag (spawning to recruitment age) | Years |
| $F'_{f,t}$ | Fully-selected fishing mortality for fleet $f$ | yr$^{-1}$ |
| $R_0$ | Unfished recruitment level | Thousands |
| $h$ | Steepness parameter in stock-recruitment | Proportion |

\pagebreak

### References

Methot, R.D. and Wetzel, C.R. (2013). Stock Synthesis: A biological and statistical framework for fish stock assessment and fishery management. *Fisheries Research*, 142: 86-99.

NMFS-OST (2025). Stock Synthesis User Manual Version 3.30.23.2. Available at: https://nmfs-ost.github.io/ss3-doc/SS330_User_Manual_release.html